---
title: "Reduced sparse monkey PK study design"
author: NAMLink
format:
  html:
    toc: true
    code-fold: false
    theme: none
    highlight-style: none
    embed-resources: true
execute:
  warning: false
  message: false
prefer-html: true
---

## Purpose

This report documents the inputs used for the reduced sparse monkey PK study
design and displays the previously completed fold-error analysis. It does not
rerun the PK model or the study-design simulations.

The analysis is predefined for atezolizumab, avelumab, nivolumab, and
pembrolizumab. Run 4 top-down estimates and relative standard errors (RSEs) are
read directly from the final monkey-model fit. Parameter uncertainty is
represented by those RSEs, and 30% interindividual variability (IIV) is
specified independently for CL, V1, Q, and V2.

```{r}
#| label: setup

rm(list = ls())

suppressPackageStartupMessages({
  library(dplyr)
  library(knitr)
  library(purrr)
  library(tibble)
})

MOLECULES <- c(
  "Atezolizumab",
  "Avelumab",
  "Nivolumab",
  "Pembrolizumab"
)
IIV_CV_PERCENT <- 30

study_dir <- if (
  basename(normalizePath(".", winslash = "/", mustWork = TRUE)) ==
    "06_StudySimulations"
) {
  normalizePath(".", winslash = "/", mustWork = TRUE)
} else {
  normalizePath("06_StudySimulations", winslash = "/", mustWork = TRUE)
}
project_dir <- normalizePath(
  file.path(study_dir, ".."),
  winslash = "/",
  mustWork = TRUE
)
run4_file <- file.path(
  project_dir,
  "04_TopDown",
  "SimsOutputs",
  "TopDownTMDD_Run4_CL_Vc_Q_Vp_fixed_TMDD_Monkey.rds"
)
figure_file <- file.path(
  study_dir,
  "SimsOutputs",
  "Figures",
  "Gate2_design_fold_error_two_vs_three_dose_by_n.png"
)
```

## Run 4 parameter inputs

```{r}
#| label: run4-parameter-inputs

run4_fits <- readRDS(run4_file)

missing_molecules <- setdiff(MOLECULES, names(run4_fits))
if (length(missing_molecules) > 0L) {
  stop(
    "Final Run 4 fits are missing for: ",
    paste(missing_molecules, collapse = ", ")
  )
}

parameter_map <- tribble(
  ~Parameter, ~Theta_name, ~Log_parameter,
  "CL",       "cl",        "lcl",
  "V1",       "vc",        "lvc",
  "Q",        "q",         "lq",
  "V2",       "vp",        "lvp"
)

extract_run4_parameters <- function(molecule) {
  fit <- run4_fits[[molecule]]$fit
  estimated_parameters <- names(fit$opt$par)
  rse <- fit$rse_percent

  parameter_map |>
    mutate(
      Molecule = molecule,
      Estimate = map_dbl(Theta_name, ~ unname(fit$theta[[.x]])),
      RSE_percent = map_dbl(
        Log_parameter,
        ~ if (.x %in% names(rse)) unname(rse[[.x]]) else NA_real_
      ),
      Run4_status = if_else(
        Log_parameter %in% estimated_parameters,
        "Estimated",
        "Fixed"
      ),
      IIV_CV_percent = IIV_CV_PERCENT
    ) |>
    select(
      Molecule,
      Parameter,
      Estimate,
      RSE_percent,
      Run4_status,
      IIV_CV_percent
    )
}

run4_parameter_inputs <- map_dfr(MOLECULES, extract_run4_parameters)

stopifnot(
  setequal(run4_parameter_inputs$Molecule, MOLECULES),
  all(is.finite(run4_parameter_inputs$Estimate)),
  all(run4_parameter_inputs$Estimate > 0),
  all(run4_parameter_inputs$IIV_CV_percent == 30)
)

run4_parameter_inputs |>
  mutate(
    Estimate = signif(Estimate, 5),
    RSE_percent = round(RSE_percent, 2)
  ) |>
  kable(
    caption = paste(
      "Final Run 4 structural parameter estimates and RSEs.",
      "Thirty percent IIV is specified separately for all four parameters."
    ),
    col.names = c(
      "Molecule",
      "Parameter",
      "Estimate",
      "RSE (%)",
      "Run 4 status",
      "IIV CV (%)"
    )
  )
```

For parameters fixed in Run 4, an RSE is not available and is shown as missing.
The 30% IIV is still applied to all four structural parameters in the completed
simulation analysis.

## Fold-error study-design figure

The figure below is reused from the completed simulations. No model simulation,
parameter sampling, or study-design resampling is performed during this
render.

```{r}
#| label: fold-error-figure
#| echo: false
#| fig-cap: "Fold-error distributions for two- and three-dose reduced sparse monkey PK study designs."
#| out-width: "100%"

if (!file.exists(figure_file)) {
  stop(
    "The completed fold-error figure is missing: ",
    figure_file,
    ". Model simulations are intentionally not rerun by this report."
  )
}

include_graphics(figure_file)
```

