---
title: "Reduced sparse monkey PK study design analysis"
author: NAMLink
format:
  html:
    toc: true
    code-fold: false
    theme: none
    highlight-style: none
    embed-resources: true
execute:
  warning: false
  message: false
prefer-html: true
---

## Purpose

This report documents the reduced sparse monkey PK study-design analysis for
AUC0-28d, Cmax, and terminal clearance. The analysis compares study arithmetic
means with full rich-population arithmetic means using the standard fold error:

`fold error = reduced-study estimate / rich-population reference`.

The prespecified acceptable interval is 0.5-2.0. Candidate group sizes are 2,
3, and 4 animals per dose group; five-animal groups are excluded from every
scenario.

The analysis is predefined for atezolizumab, avelumab, nivolumab, and
pembrolizumab. Run 4 top-down estimates and relative standard errors (RSEs) are
read directly from the final monkey-model fit. Parameter uncertainty is
represented by available RSEs, and 30% interindividual variability (IIV) is
specified independently for CL, V1, Q, and V2. A structural parameter fixed in
Run 4, and therefore missing an estimated RSE, is assigned a 40% RSE.

```{r}
#| label: setup

rm(list = ls())

suppressPackageStartupMessages({
  library(dplyr)
  library(knitr)
  library(purrr)
  library(tibble)
})

MOLECULES <- c(
  "Atezolizumab",
  "Avelumab",
  "Nivolumab",
  "Pembrolizumab"
)
IIV_CV_PERCENT <- 30
N_PER_GROUP <- 2:4
ENDPOINTS <- c("AUC0-28d", "Cmax", "Terminal clearance")

study_dir <- if (
  basename(normalizePath(".", winslash = "/", mustWork = TRUE)) ==
    "06_StudySimulations"
) {
  normalizePath(".", winslash = "/", mustWork = TRUE)
} else {
  normalizePath("06_StudySimulations", winslash = "/", mustWork = TRUE)
}
project_dir <- normalizePath(
  file.path(study_dir, ".."),
  winslash = "/",
  mustWork = TRUE
)
run4_file <- file.path(
  project_dir,
  "04_TopDown",
  "SimsOutputs",
  "TopDownTMDD_Run4_CL_Vc_Q_Vp_fixed_TMDD_Monkey.rds"
)
fold_error_figure_file <- file.path(
  study_dir,
  "SimsOutputs",
  "Figures",
  "Gate2_design_standard_fold_error_by_endpoint.png"
)
power_figure_file <- file.path(
  study_dir,
  "SimsOutputs",
  "Figures",
  "Gate2_design_power_by_endpoint.png"
)
endpoint_success_file <- file.path(
  study_dir,
  "SimsOutputs",
  "Tables",
  "Gate2_design_endpoint_success.csv"
)
```

## Run 4 parameter inputs

```{r}
#| label: run4-parameter-inputs

run4_fits <- readRDS(run4_file)

missing_molecules <- setdiff(MOLECULES, names(run4_fits))
if (length(missing_molecules) > 0L) {
  stop(
    "Final Run 4 fits are missing for: ",
    paste(missing_molecules, collapse = ", ")
  )
}

parameter_map <- tribble(
  ~Parameter, ~Theta_name, ~Log_parameter,
  "CL",       "cl",        "lcl",
  "V1",       "vc",        "lvc",
  "Q",        "q",         "lq",
  "V2",       "vp",        "lvp"
)

extract_run4_parameters <- function(molecule) {
  fit <- run4_fits[[molecule]]$fit
  estimated_parameters <- names(fit$opt$par)
  rse <- fit$rse_percent

  parameter_map |>
    mutate(
      Molecule = molecule,
      Estimate = map_dbl(Theta_name, ~ unname(fit$theta[[.x]])),
      RSE_percent = map_dbl(
        Log_parameter,
        ~ if (.x %in% names(rse)) unname(rse[[.x]]) else NA_real_
      ),
      Run4_status = if_else(
        Log_parameter %in% estimated_parameters,
        "Estimated",
        "Fixed"
      ),
      IIV_CV_percent = IIV_CV_PERCENT
    ) |>
    select(
      Molecule,
      Parameter,
      Estimate,
      RSE_percent,
      Run4_status,
      IIV_CV_percent
    )
}

run4_parameter_inputs <- map_dfr(MOLECULES, extract_run4_parameters)

stopifnot(
  setequal(run4_parameter_inputs$Molecule, MOLECULES),
  all(is.finite(run4_parameter_inputs$Estimate)),
  all(run4_parameter_inputs$Estimate > 0),
  all(run4_parameter_inputs$IIV_CV_percent == 30)
)

run4_parameter_inputs |>
  mutate(
    Estimate = signif(Estimate, 5),
    RSE_percent = round(RSE_percent, 2)
  ) |>
  kable(
    caption = paste(
      "Final Run 4 structural parameter estimates and RSEs.",
      "Thirty percent IIV is specified separately for all four parameters."
    ),
    col.names = c(
      "Molecule",
      "Parameter",
      "Estimate",
      "RSE (%)",
      "Run 4 status",
      "IIV CV (%)"
    )
  )
```

For parameters fixed in Run 4, the fitted RSE is unavailable and is shown as
missing in the input table. The simulation assigns those parameters a 40% RSE.
The independent 30% IIV is applied to all four structural parameters.

## Analysis definitions

For each simulated study replicate, animals are sampled independently within
each selected dose group. AUC0-28d, Cmax, and terminal clearance are summarized
with arithmetic means. Terminal clearance follows the project-wide PK
convention: the terminal slope is selected from consecutive post-Cmax suffixes
using the largest adjusted R-squared, `AUCinf = AUClast + Clast/lambda_z`, and
whole-animal clearance is `24 * Dose * 3.5 kg / AUCinf`.

The two-dose design uses 0.3 and 10 mg/kg, and the three-dose design uses 0.3,
3, and 10 mg/kg. A design-level endpoint result is within 2-fold only if every
selected dose group has a standard fold error between 0.5 and 2.0.

## Standard fold-error distributions

The analysis script `RunReducedSparsePKAnalysis.R` generates the simulation
outputs. The report reads those completed outputs rather than rerunning the
model during rendering.

```{r}
#| label: standard-fold-error-figure
#| echo: false
#| fig-cap: "Standard fold-error distributions for AUC0-28d, Cmax, and terminal clearance."
#| out-width: "100%"

if (!file.exists(fold_error_figure_file)) {
  stop(
    "The completed standard fold-error figure is missing: ",
    fold_error_figure_file,
    ". Run RunReducedSparsePKAnalysis.R first."
  )
}

include_graphics(fold_error_figure_file)
```

## Power-style operating characteristics

The y-axis is the fraction of simulated studies for which every selected dose
group is within 2-fold for the indicated PK endpoint.

```{r}
#| label: power-figure
#| echo: false
#| fig-cap: "Fraction of simulated studies within 2-fold by sample size per dose group."
#| out-width: "100%"

if (!file.exists(power_figure_file)) {
  stop(
    "The completed power-style figure is missing: ",
    power_figure_file,
    ". Run RunReducedSparsePKAnalysis.R first."
  )
}

include_graphics(power_figure_file)
```

## Endpoint success summary

```{r}
#| label: endpoint-success-summary

if (!file.exists(endpoint_success_file)) {
  stop(
    "The endpoint success table is missing: ",
    endpoint_success_file,
    ". Run RunReducedSparsePKAnalysis.R first."
  )
}

endpoint_success <- read.csv(
  endpoint_success_file,
  stringsAsFactors = FALSE
)

stopifnot(
  setequal(sort(unique(endpoint_success$N_per_group)), N_PER_GROUP),
  !any(endpoint_success$N_per_group == 5),
  setequal(unique(endpoint_success$Endpoint), ENDPOINTS)
)

endpoint_success |>
  mutate(
    Fraction_within_2fold = round(Fraction_within_2fold, 3)
  ) |>
  arrange(Drug, Endpoint, Design, N_per_group) |>
  kable(
    caption = paste(
      "Fraction of simulated studies within 2-fold for every selected dose",
      "group, by endpoint, design, and group size."
    ),
    col.names = c(
      "Drug",
      "Design",
      "Animals per dose",
      "Total animals",
      "Endpoint",
      "Fraction within 2-fold"
    )
  )
```
