---
title: "Top-down QSS TMDD fits for monkey IV PK"
format: html
---

# Set up

```{r}
#| label: setup
#| echo: true
#| warning: false
#| message: false

rm(list = ls())

library(tidyverse)
library(data.table)
library(scales)
library(openxlsx)
library(knitr)
library(kableExtra)
library(nlmixr2)
library(rxode2)
library(nlmixr2lib)

datadir <- if (dir.exists("../../01_Data")) "../../01_Data" else "01_Data"
outdir <- if (dir.exists("../SimsOutputs") || dir.exists("../Scripts")) "../SimsOutputs/" else "04_TopDown/SimsOutputs/"
dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
figures_dir <- file.path(outdir, "Figures")
dir.create(figures_dir, recursive = TRUE, showWarnings = FALSE)

AUC_END_H <- 28 * 24
FORCE_RERUN <- FALSE

normalize_names <- function(x) {
  x %>%
    str_replace_all("\\.+", ".") %>%
    str_replace_all("\\s+", ".") %>%
    str_replace_all("\\[", "") %>%
    str_replace_all("\\]", "") %>%
    str_replace_all("\\(", "") %>%
    str_replace_all("\\)", "")
}

normalize_species <- function(x) {
  case_when(
    str_to_lower(x) %in% c("nhp", "monkey", "cynomolgus monkey") ~ "Monkey",
    TRUE ~ str_to_title(x)
  )
}

unit_to_hour <- function(time, unit) {
  unit_l <- str_to_lower(str_trim(unit))
  case_when(
    unit_l %in% c("h", "hr", "hrs", "hour", "hours") ~ as.numeric(time),
    unit_l %in% c("min", "mins", "minute", "minutes") ~ as.numeric(time) / 60,
    unit_l %in% c("day", "days", "d") ~ as.numeric(time) * 24,
    TRUE ~ as.numeric(time)
  )
}

conc_ugml_to_nm <- function(conc_ugml, mw_g_mol) {
  as.numeric(conc_ugml) * 1e6 / as.numeric(mw_g_mol)
}

conc_nm_to_ugml <- function(conc_nm, mw_g_mol) {
  as.numeric(conc_nm) * as.numeric(mw_g_mol) / 1e6
}

dose_mgkg_to_nmolkg <- function(dose_mgkg, mw_g_mol) {
  as.numeric(dose_mgkg) * 1e6 / as.numeric(mw_g_mol)
}

convert_rate_to_h <- function(value, unit) {
  unit_l <- str_to_lower(str_trim(unit))
  case_when(
    unit_l %in% c("1/min", "min^-1", "1/minute") ~ as.numeric(value) * 60,
    TRUE ~ as.numeric(value)
  )
}

convert_kd_to_nm <- function(value, unit) {
  unit_l <- str_to_lower(str_trim(unit))
  unit_l <- str_replace_all(unit_l, "[^[:ascii:]]", "u")
  case_when(
    unit_l %in% c("um", "umol/l", "umol/litre", "umol/liter", "umol/l") ~ as.numeric(value) * 1000,
    unit_l %in% c("nm", "nmol/l", "nmol/litre", "nmol/liter") ~ as.numeric(value),
    TRUE ~ as.numeric(value)
  )
}

bounded_start <- function(x, default, lower, upper) {
  x <- suppressWarnings(as.numeric(x))
  x[!is.finite(x) | x <= 0] <- default
  pmin(pmax(x, lower), upper)
}

run_result_path <- function(run_name) {
  file.path(outdir, paste0("TopDownTMDD_", run_name, "_Monkey.rds"))
}

fit_or_read_run <- function(run_name, fit_expr, force = FORCE_RERUN) {
  result_file <- run_result_path(run_name)
  if (file.exists(result_file) && !isTRUE(force)) {
    message("Reading cached ", run_name, " results from ", result_file)
    return(readRDS(result_file))
  }

  result <- force(fit_expr)
  saveRDS(result, result_file)
  message("Saved ", run_name, " results to ", result_file)
  result
}

terminal_lambda_z <- function(time_h, conc) {
  df <- tibble(time_h = as.numeric(time_h), conc = as.numeric(conc)) %>%
    filter(is.finite(time_h), is.finite(conc), conc > 0) %>%
    arrange(time_h)

  if (nrow(df) < 3) return(NA_real_)

  tail_df <- tail(df, min(4, nrow(df)))
  fit <- tryCatch(lm(log(conc) ~ time_h, data = tail_df), error = function(e) NULL)
  if (is.null(fit)) return(NA_real_)

  lambda <- -unname(coef(fit)[["time_h"]])
  if (!is.finite(lambda) || lambda <= 0) NA_real_ else lambda
}

trapz_auc <- function(time, conc, t_end = Inf) {
  keep <- !is.na(time) & !is.na(conc) & conc > 0 & time <= t_end
  t <- as.numeric(time[keep])
  c <- as.numeric(conc[keep])
  if (length(t) < 2) return(NA_real_)

  ord <- order(t)
  t <- t[ord]
  c <- c[ord]
  dt <- diff(t)
  c1 <- c[-length(c)]
  c2 <- c[-1]
  use_log <- c1 != c2 & c1 > 0 & c2 > 0
  sum(ifelse(use_log, dt * (c1 - c2) / log(c1 / c2), dt * (c1 + c2) / 2))
}

fmt_fe <- function(x) {
  cell_spec(
    ifelse(is.na(x), "", formatC(x, format = "f", digits = 2)),
    background = ifelse(!is.na(x) & x >= 0.5 & x <= 2, "#d4edda", "#f8d7da"),
    color = ifelse(!is.na(x) & x >= 0.5 & x <= 2, "#155724", "#721c24"),
    bold = TRUE
  )
}

safe_ofv <- function(fit) {
  candidates <- list(
    quote(fit$objf),
    quote(fit$objective),
    quote(fit$OBJF),
    quote(fit$ofv)
  )
  for (expr in candidates) {
    val <- tryCatch(eval(expr), error = function(e) NULL)
    if (!is.null(val) && length(val) > 0 && is.finite(as.numeric(val)[1])) {
      return(as.numeric(val)[1])
    }
  }
  NA_real_
}

if (.Platform$OS.type == "windows" && identical(unname(Sys.which("make")), "")) {
  warning(
    "Rtools/make was not found on PATH. nlmixr2/rxode2 model compilation may fail; ",
    "run nlmixr2::nlmixr2CheckInstall() if Run1 reports rxode2 compile errors.",
    call. = FALSE
  )
}
```

# Read data and parameters

```{r}
#| label: read-data

obs_raw <- read.csv(
  file.path(datadir, "Monkey", "mAb_data_clean.csv"),
  stringsAsFactors = FALSE,
  check.names = FALSE
)
names(obs_raw) <- normalize_names(names(obs_raw))

param_raw <- read.xlsx(file.path(datadir, "Monkey", "Parameters.xlsx"), sheet = "Included")
names(param_raw) <- normalize_names(names(param_raw))
included_molecules <- unique(param_raw$Drug.Name)

obs_points <- obs_raw %>%
  transmute(
    Study.Id = as.character(.data[["Study.Id"]]),
    Subject.Id = coalesce(
      na_if(as.character(.data[["Subject.Id"]]), "."),
      na_if(as.character(.data[["Group.Id"]]), "."),
      as.character(.data[["Study.Id"]])
    ),
    Species = normalize_species(.data[["Species"]]),
    Dose = as.numeric(.data[["Dose"]]),
    Molecule = .data[["Molecule"]],
    MW = as.numeric(.data[["Molecular.Weight"]]),
    Time_h = unit_to_hour(.data[["Time"]], .data[["Time.unit"]]),
    Conc_ugmL = as.numeric(.data[["Measurement"]]),
    Conc_unit = .data[["Measurement.unit"]],
    Route = .data[["Route"]]
  ) %>%
  filter(
    Species == "Monkey",
    Molecule %in% included_molecules,
    Route %in% c("IV", "IV Bolus"),
    str_to_lower(Conc_unit) %in% c("ug/ml", "microgram/ml", "micrograms/ml"),
    !is.na(Time_h),
    !is.na(Conc_ugmL),
    Conc_ugmL > 0
  ) %>%
  mutate(
    Profile = paste(Molecule, Dose, Subject.Id, Study.Id, sep = "_"),
    ID = as.integer(factor(Profile)),
    DV = conc_ugml_to_nm(Conc_ugmL, MW)
  )

obs <- obs_points %>%
  group_by(Molecule, Species, Dose, MW, Time_h) %>%
  summarise(
    Conc_ugmL = median(Conc_ugmL, na.rm = TRUE),
    DV = median(DV, na.rm = TRUE),
    n_points = n(),
    .groups = "drop"
  ) %>%
  arrange(Molecule, Dose, Time_h) %>%
  mutate(
    Study.Id = "Median_by_molecule_dose_time",
    Subject.Id = paste(Molecule, Dose, "median", sep = "_"),
    Route = "IV",
    Profile = paste(Molecule, Dose, sep = "_"),
    ID = as.integer(factor(Profile))
  ) %>%
  select(
    Study.Id, Subject.Id, Species, Dose, Molecule, MW, Time_h,
    Conc_ugmL, Route, Profile, ID, DV, n_points
  )

stopifnot(length(unique(obs$Molecule)) == 8)

params <- param_raw %>%
  transmute(
    Molecule = .data[["Drug.Name"]],
    Target = .data[["Target.Name"]],
    Parameter = .data[["Parameter.Name.in.PK-Sim"]],
    Value = as.numeric(.data[["Value.from.Source"]]),
    Unit = .data[["Unit"]],
    Source = .data[["Source"]]
  ) %>%
  filter(Molecule %in% unique(obs$Molecule)) %>%
  mutate(
    parameter_key = case_when(
      str_detect(str_to_lower(Parameter), "^kd$") ~ "kd_nm",
      str_to_lower(Parameter) == "koff" ~ "koff_h",
      str_to_lower(Parameter) == "kint" ~ "kint_h",
      str_to_lower(Parameter) == "kdeg" ~ "kdeg_h",
      str_to_lower(Parameter) == "reference concentration" ~ "ref_nm",
      TRUE ~ NA_character_
    ),
    unit_for_conversion = case_when(
      parameter_key == "koff_h" &
        !coalesce(str_detect(str_to_lower(Unit), "1/|min|h|hour"), FALSE) ~ "1/min",
      TRUE ~ Unit
    ),
    value_std = case_when(
      parameter_key == "kd_nm" ~ convert_kd_to_nm(Value, Unit),
      parameter_key == "ref_nm" ~ convert_kd_to_nm(Value, Unit),
      parameter_key == "koff_h" ~ convert_rate_to_h(Value, unit_for_conversion),
      parameter_key %in% c("kint_h", "kdeg_h") ~ convert_rate_to_h(Value, Unit),
      TRUE ~ NA_real_
    )
  ) %>%
  filter(!is.na(parameter_key)) %>%
  select(Molecule, Target, parameter_key, value_std) %>%
  pivot_wider(names_from = parameter_key, values_from = value_std) %>%
  mutate(
    kon_inv_nmh = koff_h / kd_nm,
    kss_qss_nm = (koff_h + kint_h) / kon_inv_nmh,
    MW = map_dbl(Molecule, ~ unique(obs$MW[obs$Molecule == .x])[1])
  )

unit_audit <- param_raw %>%
  transmute(
    Molecule = .data[["Drug.Name"]],
    Parameter = .data[["Parameter.Name.in.PK-Sim"]],
    Value = as.numeric(.data[["Value.from.Source"]]),
    Unit = .data[["Unit"]]
  ) %>%
  filter(
    Molecule %in% unique(obs$Molecule),
    Parameter %in% c("Kd", "koff", "kint", "kdeg")
  ) %>%
  left_join(
    params %>%
      select(Molecule, kd_nm, koff_h, kint_h, kdeg_h, kon_inv_nmh, kss_qss_nm),
    by = "Molecule"
  )

knitr::kable(params, digits = 4, caption = "Fixed monkey TMDD inputs used by the top-down fits")
knitr::kable(unit_audit, digits = 4, caption = "Unit conversion audit for TMDD input parameters")

obs %>%
  count(Molecule, Dose, ID, name = "n_time_points") %>%
  arrange(Molecule, Dose) %>%
  knitr::kable(caption = "Median monkey single-dose IV profiles used for fitting after resetting IDs")

obs %>%
  distinct(Molecule, Dose, ID, MW) %>%
  transmute(
    Molecule,
    Dose_mg_kg = Dose,
    ID,
    Dose_events_per_ID = 1L,
    Dose_time_h = 0,
    Dose_nmol_kg = dose_mgkg_to_nmolkg(Dose, MW)
  ) %>%
  arrange(Molecule, Dose_mg_kg) %>%
  knitr::kable(digits = 4, caption = "Single-dose IV event audit: one dose at time zero per molecule-dose profile")

profile_starts <- obs %>%
  group_by(Molecule, Dose, MW) %>%
  summarise(
    dose_nmolkg = dose_mgkg_to_nmolkg(first(Dose), first(MW)),
    cmax_nm = max(DV, na.rm = TRUE),
    lambda_z_h = terminal_lambda_z(Time_h, DV),
    vc_lkg = dose_nmolkg / cmax_nm,
    cl_lhkg = lambda_z_h * vc_lkg,
    .groups = "drop"
  )

initial_estimates <- profile_starts %>%
  group_by(Molecule) %>%
  summarise(
    lcl_init = log(bounded_start(median(cl_lhkg, na.rm = TRUE), 0.01, 1e-4, 0.2)),
    lvc_init = log(bounded_start(median(vc_lkg, na.rm = TRUE), 0.05, 0.005, 0.5)),
    lq_init = log(bounded_start(median(cl_lhkg, na.rm = TRUE) * 0.5, 0.005, 1e-5, 0.2)),
    lvp_init = log(bounded_start(median(vc_lkg, na.rm = TRUE) * 1.5, 0.08, 0.005, 1.0)),
    add_err_init = bounded_start(0.05 * median(cmax_nm, na.rm = TRUE), 0.5, 1e-4, 1000),
    .groups = "drop"
  ) %>%
  left_join(params %>% select(Molecule, ref_nm, kdeg_h), by = "Molecule") %>%
  mutate(
    lr0_init = log(bounded_start(ref_nm, 0.1, 1e-5, 1e4)),
    lkdeg_init = log(bounded_start(kdeg_h, 0.03, 1e-5, 10)),
    prop_err_init = 0.12,
    CL_init_L_h_kg = exp(lcl_init),
    Vc_init_L_kg = exp(lvc_init),
    Q_init_L_h_kg = exp(lq_init),
    Vp_init_L_kg = exp(lvp_init),
    Rtot_init_nM = exp(lr0_init),
    kdeg_init_h = exp(lkdeg_init)
  )

initial_estimates %>%
  select(
    Molecule, CL_init_L_h_kg, Vc_init_L_kg, Q_init_L_h_kg, Vp_init_L_kg,
    Rtot_init_nM, kdeg_init_h, prop_err_init, add_err_init
  ) %>%
  knitr::kable(digits = 4, caption = "Data-driven initial estimates used by nlmixr2")
```

# TMDD QSS model helpers

The QSS model follows the two-compartment QSS TMDD form used in the Pumas and Monolix examples: central drug is treated as total ligand, total target is an ODE state, and free ligand/free target/complex are computed algebraically. Concentration-bearing quantities are in nM, time is in hours, amount is nmol/kg, and volume/clearance are weight-normalized.

For this top-down exercise, the target baseline is estimated as `Rtot`. Target synthesis is handled explicitly as:

`ksyn = Rtot * kdeg_target`.

The QSS binding constant is fixed from the in vitro kinetic inputs:

`Kss = (koff + kint) / kon`, where `kon = koff / KD`.

Run 1 fixes `kdeg_target` to the parameter workbook value and estimates `Rtot`. Run 2 estimates both `Rtot` and `kdeg_target`. Run 3 fixes `Rtot` to the value provided in the input dataset and estimates `CL` and `kint`.
Unless otherwise stated for a run, the structural PK terms are fixed to data-derived monkey profile starts so the top-down step does not try to solve a poorly identified joint PK/TMDD problem from median-only profiles.

```{r}
#| label: model-helpers

make_fit_data <- function(dat) {
  dat_obs <- dat %>%
    arrange(ID, Time_h) %>%
    transmute(
      ID,
      TIME = Time_h,
      DV,
      AMT = NA_real_,
      EVID = 0,
      CMT = "centr",
      Molecule,
      Dose,
      MW
    )

  dose_records <- dat %>%
    distinct(ID, Molecule, Dose, MW) %>%
    transmute(
      ID,
      TIME = 0,
      DV = NA_real_,
      AMT = dose_mgkg_to_nmolkg(Dose, MW),
      EVID = 1,
      CMT = "centr",
      Molecule,
      Dose,
      MW
    )

  bind_rows(dose_records, dat_obs) %>%
    arrange(ID, TIME, desc(EVID)) %>%
    as.data.frame()
}

build_qss_model <- function(kd_nm, kon_inv_nmh, kint_h, kdeg_h = NULL, kss_nm = NULL, estimate_kdeg = FALSE, init = NULL) {
  kd_nm <- signif(as.numeric(kd_nm), 8)
  kon_inv_nmh <- signif(as.numeric(kon_inv_nmh), 8)
  kint_h <- signif(as.numeric(kint_h), 8)
  kdeg_h <- signif(as.numeric(kdeg_h), 8)
  kss_nm <- signif(as.numeric(kss_nm), 8)
  default_init <- list(
    lcl_init = log(0.015),
    lvc_init = log(0.045),
    lq_init = log(0.020),
    lvp_init = log(0.055),
    lr0_init = log(0.1),
    lkdeg_init = log(max(kdeg_h, 1e-4)),
    prop_err_init = 0.09,
    add_err_init = 10
  )
  init <- modifyList(default_init, as.list(init))
  init <- lapply(init, function(x) signif(as.numeric(x)[1], 8))

  if (estimate_kdeg) {
    model_txt <- glue::glue("
      function() {{
        ini({{
          lcl <- fixed({init$lcl_init})
          lvc <- fixed({init$lvc_init})
          lq <- fixed({init$lq_init})
          lvp <- fixed({init$lvp_init})
          lr0 <- c(log(1e-8), {init$lr0_init}, log(1e4))
          lkdeg <- c(log(1e-4), {init$lkdeg_init}, log(2))
          prop.err <- c(1e-4, {init$prop_err_init}, 1.5)
          add.err <- c(1e-4, {init$add_err_init}, 1e4)
        }})
        model({{
          cl <- exp(lcl)
          vc <- exp(lvc)
          q <- exp(lq)
          vp <- exp(lvp)
          r0 <- exp(lr0)
          kdeg <- exp(lkdeg)
          kd <- {kd_nm}
          kon <- {kon_inv_nmh}
          kint <- {kint_h}
          kss <- {kss_nm}
          ksyn <- r0 * kdeg

          Ltot <- centr / vc
          Cp <- peri / vp
          free_drug <- 0.5 * ((Ltot - total_target - kss) + sqrt((Ltot - total_target - kss)^2 + 4 * kss * Ltot + 1e-12))
          complex <- Ltot - free_drug
          free_target <- total_target - complex

          d/dt(centr) = -cl * free_drug - q * free_drug + q * Cp - kint * complex * vc
          d/dt(peri) = q * free_drug - q * Cp
          d/dt(total_target) = ksyn - kdeg * free_target - kint * complex
          total_target(0) = r0

          Ltot ~ add(add.err) + prop(prop.err)
        }})
      }}
    ")
  } else {
    model_txt <- glue::glue("
      function() {{
        ini({{
          lcl <- fixed({init$lcl_init})
          lvc <- fixed({init$lvc_init})
          lq <- fixed({init$lq_init})
          lvp <- fixed({init$lvp_init})
          lr0 <- c(log(1e-8), {init$lr0_init}, log(1e4))
          prop.err <- c(1e-4, {init$prop_err_init}, 1.5)
          add.err <- c(1e-4, {init$add_err_init}, 1e4)
        }})
        model({{
          cl <- exp(lcl)
          vc <- exp(lvc)
          q <- exp(lq)
          vp <- exp(lvp)
          r0 <- exp(lr0)
          kdeg <- {kdeg_h}
          kd <- {kd_nm}
          kon <- {kon_inv_nmh}
          kint <- {kint_h}
          kss <- {kss_nm}
          ksyn <- r0 * kdeg

          Ltot <- centr / vc
          Cp <- peri / vp
          free_drug <- 0.5 * ((Ltot - total_target - kss) + sqrt((Ltot - total_target - kss)^2 + 4 * kss * Ltot + 1e-12))
          complex <- Ltot - free_drug
          free_target <- total_target - complex

          d/dt(centr) = -cl * free_drug - q * free_drug + q * Cp - kint * complex * vc
          d/dt(peri) = q * free_drug - q * Cp
          d/dt(total_target) = ksyn - kdeg * free_target - kint * complex
          total_target(0) = r0

          Ltot ~ add(add.err) + prop(prop.err)
        }})
      }}
    ")
  }

  eval(parse(text = model_txt))
}

extract_fit_table <- function(fit, run_name, molecule, param_row) {
  fixed <- tryCatch(fixef(fit), error = function(e) NULL)
  if (is.null(fixed) || length(fixed) == 0) {
    fixed <- tryCatch(coef(fit)$fixed, error = function(e) NULL)
  }
  if (is.null(fixed) || length(fixed) == 0) {
    par_fixed <- tryCatch(fit$parFixed, error = function(e) NULL)
    if (is.data.frame(par_fixed) && "Est." %in% names(par_fixed)) {
      fixed <- setNames(suppressWarnings(as.numeric(par_fixed[["Est."]])), rownames(par_fixed))
    }
  }
  if (is.null(fixed) || length(fixed) == 0) stop("Could not extract fixed effects from fit object")

  theta <- tibble(
    parameter = names(fixed),
    estimate = as.numeric(fixed)
  )

  get_est <- function(parameter) {
    val <- theta$estimate[theta$parameter == parameter]
    if (length(val) == 0) NA_real_ else as.numeric(val[1])
  }

  rtot <- exp(get_est("lr0"))
  kdeg <- if ("lkdeg" %in% theta$parameter) exp(get_est("lkdeg")) else param_row$kdeg_h
  kss <- param_row$kss_qss_nm

  tibble(
    Run = run_name,
    Molecule = molecule,
    OFV = safe_ofv(fit),
    Rtot_nM = rtot,
    kdeg_h = kdeg,
    Kss_nM = kss,
    CL_L_h_kg = exp(get_est("lcl")),
    Vc_L_kg = exp(get_est("lvc")),
    Q_L_h_kg = exp(get_est("lq")),
    Vp_L_kg = exp(get_est("lvp")),
    Fit_Status = "success",
    Fit_Error = NA_character_
  )
}

extract_predictions <- function(fit, run_name, molecule, mw) {
  pred_df <- as.data.frame(fit)
  names(pred_df) <- toupper(names(pred_df))
  pred_col <- intersect(c("IPRED", "PRED", "LTOT", "CC", "CTOT", "CP"), names(pred_df))[1]
  if (is.na(pred_col)) {
    stop("No prediction column found for ", molecule, " ", run_name)
  }

  if ("EVID" %in% names(pred_df)) {
    pred_df <- filter(pred_df, EVID == 0 | is.na(EVID))
  }

  pred_df %>%
    transmute(
      Run = run_name,
      ID = as.integer(ID),
      TIME = as.numeric(TIME),
      DV = as.numeric(DV),
      Pred_nM = as.numeric(.data[[pred_col]]),
      Pred_ugmL = conc_nm_to_ugml(Pred_nM, mw),
      Molecule = molecule
    )
}

qss_rxode_model <- rxode2::rxode({
  Ltot <- centr / vc
  Cp <- peri / vp
  free_drug <- 0.5 * ((Ltot - total_target - kss) + sqrt((Ltot - total_target - kss)^2 + 4 * kss * Ltot + 1e-12))
  complex <- Ltot - free_drug
  free_target <- total_target - complex

  d/dt(centr) = -cl * free_drug - q * free_drug + q * Cp - kint * complex * vc
  d/dt(peri) = q * free_drug - q * Cp
  d/dt(total_target) = ksyn - kdeg * free_target - kint * complex
})

simulate_qss_profiles <- function(dat_mol, theta, run_name) {
  map_dfr(unique(dat_mol$ID), function(profile_id) {
    dat_id <- dat_mol %>%
      filter(ID == profile_id) %>%
      arrange(Time_h)

    dose_nmolkg <- dose_mgkg_to_nmolkg(dat_id$Dose[1], dat_id$MW[1])
    event_table <- data.frame(
      id = profile_id,
      time = c(0, dat_id$Time_h),
      amt = c(dose_nmolkg, rep(NA_real_, nrow(dat_id))),
      evid = c(1, rep(0, nrow(dat_id))),
      cmt = "centr"
    )

    sol <- rxode2::rxSolve(
      qss_rxode_model,
      params = theta,
      events = event_table,
      inits = c(centr = 0, peri = 0, total_target = theta[["r0"]]),
      rtol = 1e-6,
      atol = 1e-9,
      hmin = 1e-12,
      maxsteps = 100000,
      returnType = "data.frame"
    ) %>%
      as_tibble() %>%
      transmute(
        ID = as.integer(profile_id),
        TIME = as.numeric(time),
        Pred_nM = pmax(as.numeric(Ltot), 0)
      )

    dat_id %>%
      transmute(
        Run = run_name,
        ID,
        TIME = Time_h,
        DV,
        Molecule,
        MW
      ) %>%
      left_join(sol, by = c("ID", "TIME")) %>%
      mutate(
        Pred_ugmL = conc_nm_to_ugml(Pred_nM, MW)
      ) %>%
      select(Run, ID, TIME, DV, Pred_nM, Pred_ugmL, Molecule)
  })
}

qss_objective <- function(log_par, dat_mol, par_mol, init_mol, run_name, estimate_kdeg = FALSE, estimate_cl = FALSE, estimate_kint = FALSE, fixed_rtot = FALSE) {
  cl <- if (estimate_cl) exp(log_par[["lcl"]]) else init_mol$CL_init_L_h_kg
  r0 <- if (fixed_rtot) init_mol$Rtot_init_nM else exp(log_par[["lr0"]])
  kdeg <- if (estimate_kdeg) exp(log_par[["lkdeg"]]) else par_mol$kdeg_h
  kint <- if (estimate_kint) exp(log_par[["lkint"]]) else par_mol$kint_h
  prop_err <- exp(log_par[["lprop"]])
  add_err <- exp(log_par[["ladd"]])

  theta <- c(
    cl = cl,
    vc = init_mol$Vc_init_L_kg,
    q = init_mol$Q_init_L_h_kg,
    vp = init_mol$Vp_init_L_kg,
    r0 = r0,
    kdeg = kdeg,
    kint = kint,
    kss = (par_mol$koff_h + kint) / par_mol$kon_inv_nmh,
    ksyn = r0 * kdeg
  )

  pred <- tryCatch(
    simulate_qss_profiles(dat_mol, theta, run_name),
    error = function(e) NULL
  )
  if (is.null(pred) || any(!is.finite(pred$Pred_nM))) return(1e100)

  sd <- pmax(add_err + prop_err * pmax(pred$Pred_nM, 0), 1e-8)
  residual <- (pred$DV - pred$Pred_nM) / sd
  sum(log(2 * pi * sd^2) + residual^2)
}

fit_qss_bobyqa <- function(dat_mol, par_mol, init_mol, run_name, estimate_kdeg = FALSE, estimate_cl = FALSE, estimate_kint = FALSE, fixed_rtot = FALSE) {
  start <- c()
  lower <- c()
  upper <- c()

  if (estimate_cl) {
    start <- c(start, lcl = init_mol$lcl_init)
    lower <- c(lower, lcl = log(1e-4))
    upper <- c(upper, lcl = log(0.5))
  }

  if (!fixed_rtot) {
    start <- c(start, lr0 = init_mol$lr0_init)
    lower <- c(lower, lr0 = log(1e-8))
    upper <- c(upper, lr0 = log(1e4))
  }

  if (estimate_kdeg) {
    start <- c(start, lkdeg = init_mol$lkdeg_init)
    lower <- c(lower, lkdeg = log(1e-4))
    upper <- c(upper, lkdeg = log(2))
  }

  if (estimate_kint) {
    start <- c(start, lkint = log(bounded_start(par_mol$kint_h, 0.03, 1e-5, 10)))
    lower <- c(lower, lkint = log(1e-5))
    upper <- c(upper, lkint = log(10))
  }

  start <- c(start, lprop = log(init_mol$prop_err_init), ladd = log(init_mol$add_err_init))
  lower <- c(lower, lprop = log(1e-4), ladd = log(1e-4))
  upper <- c(upper, lprop = log(1.5), ladd = log(1e4))

  start <- pmin(pmax(start, lower + 1e-8), upper - 1e-8)
  par_names <- names(start)
  objective_named <- function(par, ...) {
    names(par) <- par_names
    qss_objective(par, ...)
  }

  opt <- minqa::bobyqa(
    par = start,
    fn = objective_named,
    lower = lower,
    upper = upper,
    control = list(maxfun = 1200, rhobeg = 0.2, rhoend = 1e-6),
    dat_mol = dat_mol,
    par_mol = par_mol,
    init_mol = init_mol,
    run_name = run_name,
    estimate_kdeg = estimate_kdeg,
    estimate_cl = estimate_cl,
    estimate_kint = estimate_kint,
    fixed_rtot = fixed_rtot
  )
  opt$par <- setNames(opt$par, par_names)

  cl <- if (estimate_cl) exp(opt$par[["lcl"]]) else init_mol$CL_init_L_h_kg
  r0 <- if (fixed_rtot) init_mol$Rtot_init_nM else exp(opt$par[["lr0"]])
  kdeg <- if (estimate_kdeg) exp(opt$par[["lkdeg"]]) else par_mol$kdeg_h
  kint <- if (estimate_kint) exp(opt$par[["lkint"]]) else par_mol$kint_h
  theta <- c(
    cl = cl,
    vc = init_mol$Vc_init_L_kg,
    q = init_mol$Q_init_L_h_kg,
    vp = init_mol$Vp_init_L_kg,
    r0 = r0,
    kdeg = kdeg,
    kint = kint,
    kss = (par_mol$koff_h + kint) / par_mol$kon_inv_nmh,
    ksyn = r0 * kdeg
  )

  list(
    opt = opt,
    theta = theta,
    predictions = simulate_qss_profiles(dat_mol, theta, run_name)
  )
}

fit_one_molecule <- function(molecule, run_name, estimate_kdeg = FALSE, estimate_cl = FALSE, estimate_kint = FALSE, fixed_rtot = FALSE) {
  dat_mol <- filter(obs, Molecule == molecule)
  par_mol <- filter(params, Molecule == molecule)
  init_mol <- filter(initial_estimates, Molecule == molecule)

  message(
    "Starting ", molecule, " ", run_name,
    ": CL=", signif(init_mol$CL_init_L_h_kg, 3),
    " L/h/kg, Vc=", signif(init_mol$Vc_init_L_kg, 3),
    " L/kg, Rtot=", signif(init_mol$Rtot_init_nM, 3),
    " nM"
  )

  fit <- fit_qss_bobyqa(dat_mol, par_mol, init_mol, run_name, estimate_kdeg, estimate_cl, estimate_kint, fixed_rtot)

  list(
    fit = fit,
    summary = tibble(
      Run = run_name,
      Molecule = molecule,
      OFV = fit$opt$fval,
      Rtot_nM = fit$theta[["r0"]],
      kdeg_h = fit$theta[["kdeg"]],
      kint_h = fit$theta[["kint"]],
      Kss_nM = fit$theta[["kss"]],
      CL_L_h_kg = fit$theta[["cl"]],
      Vc_L_kg = fit$theta[["vc"]],
      Q_L_h_kg = fit$theta[["q"]],
      Vp_L_kg = fit$theta[["vp"]],
      Fit_Status = ifelse(fit$opt$ierr == 0, "success", paste0("bobyqa_ierr_", fit$opt$ierr)),
      Fit_Error = ifelse(fit$opt$ierr == 0, NA_character_, fit$opt$msg)
    ),
    predictions = fit$predictions
  )
}

safe_fit_one_molecule <- function(molecule, run_name, estimate_kdeg = FALSE, estimate_cl = FALSE, estimate_kint = FALSE, fixed_rtot = FALSE) {
  tryCatch(
    fit_one_molecule(molecule, run_name, estimate_kdeg, estimate_cl, estimate_kint, fixed_rtot),
    error = function(e) {
      warning(run_name, " failed for ", molecule, ": ", conditionMessage(e), call. = FALSE)
      list(
        fit = NULL,
        summary = tibble(
          Run = run_name,
          Molecule = molecule,
          OFV = NA_real_,
          Rtot_nM = NA_real_,
          kdeg_h = NA_real_,
          kint_h = NA_real_,
          Kss_nM = NA_real_,
          CL_L_h_kg = NA_real_,
          Vc_L_kg = NA_real_,
          Q_L_h_kg = NA_real_,
          Vp_L_kg = NA_real_,
          Fit_Status = "failed",
          Fit_Error = conditionMessage(e)
        ),
        predictions = tibble(
          Run = character(),
          ID = integer(),
          TIME = numeric(),
          DV = numeric(),
          Pred_nM = numeric(),
          Pred_ugmL = numeric(),
          Molecule = character()
        )
      )
    }
  )
}
```

# TMDD QSS Fits: Run1

```{r}
#| label: Run1
#| warning: false
#| message: false

molecules <- sort(unique(obs$Molecule))

run1_fits <- fit_or_read_run(
  "Run1_Rtot",
  set_names(molecules) %>%
    map(~ safe_fit_one_molecule(.x, run_name = "Run1_Rtot", estimate_kdeg = FALSE))
)

run1_fit_summary <- map_dfr(run1_fits, "summary")
run1_predictions <- map_dfr(run1_fits, "predictions")

run1_fit_summary %>%
  select(Run, Molecule, Fit_Status, OFV, Rtot_nM, kdeg_h, kint_h, Kss_nM, CL_L_h_kg, Fit_Error) %>%
  arrange(Molecule) %>%
  kable(digits = 4, caption = "Run 1: QSS TMDD fits estimating Rtot")
```

# TMDD QSS Fits: Run2

```{r}
#| label: Run2
#| warning: false
#| message: false

run2_fits <- fit_or_read_run(
  "Run2_Rtot_kdeg",
  set_names(molecules) %>%
    map(~ safe_fit_one_molecule(.x, run_name = "Run2_Rtot_kdeg", estimate_kdeg = TRUE))
)

run2_fit_summary <- map_dfr(run2_fits, "summary")
run2_predictions <- map_dfr(run2_fits, "predictions")

run2_fit_summary %>%
  select(Run, Molecule, Fit_Status, OFV, Rtot_nM, kdeg_h, kint_h, Kss_nM, CL_L_h_kg, Fit_Error) %>%
  arrange(Molecule) %>%
  kable(digits = 4, caption = "Run 2: QSS TMDD fits estimating Rtot and kdeg")
```

# TMDD QSS Fits: Run3

```{r}
#| label: Run3
#| warning: false
#| message: false

run3_fits <- fit_or_read_run(
  "Run3_CL_kint_fixed_Rtot",
  set_names(molecules) %>%
    map(~ safe_fit_one_molecule(
      .x,
      run_name = "Run3_CL_kint_fixed_Rtot",
      estimate_cl = TRUE,
      estimate_kint = TRUE,
      fixed_rtot = TRUE
    ))
)

run3_fit_summary <- map_dfr(run3_fits, "summary")
run3_predictions <- map_dfr(run3_fits, "predictions")

run3_fit_summary %>%
  select(Run, Molecule, Fit_Status, OFV, Rtot_nM, kdeg_h, kint_h, Kss_nM, CL_L_h_kg, Fit_Error) %>%
  arrange(Molecule) %>%
  kable(digits = 4, caption = "Run 3: QSS TMDD fits estimating CL and kint with fixed dataset Rtot")
```

# Process outputs

```{r}
#| label: process-outputs

fit_summary <- bind_rows(run1_fit_summary, run3_fit_summary) %>%
  arrange(Molecule, Run)

predictions <- bind_rows(run1_predictions, run3_predictions)

obs_metrics <- obs %>%
  group_by(Molecule, Dose, ID) %>%
  summarise(
    Obs_Cmax = max(Conc_ugmL, na.rm = TRUE),
    Obs_AUC = trapz_auc(Time_h, Conc_ugmL, t_end = AUC_END_H),
    .groups = "drop"
  ) %>%
  group_by(Molecule, Dose) %>%
  summarise(
    Obs_Cmax = mean(Obs_Cmax, na.rm = TRUE),
    Obs_AUC = mean(Obs_AUC, na.rm = TRUE),
    .groups = "drop"
  )

pred_metrics <- predictions %>%
  left_join(obs %>% distinct(ID, Molecule, Dose), by = c("ID", "Molecule")) %>%
  group_by(Run, Molecule, Dose, ID) %>%
  summarise(
    Pred_Cmax = max(Pred_ugmL, na.rm = TRUE),
    Pred_AUC = trapz_auc(TIME, Pred_ugmL, t_end = AUC_END_H),
    .groups = "drop"
  ) %>%
  group_by(Run, Molecule, Dose) %>%
  summarise(
    Pred_Cmax = mean(Pred_Cmax, na.rm = TRUE),
    Pred_AUC = mean(Pred_AUC, na.rm = TRUE),
    .groups = "drop"
  )

metrics_tbl <- pred_metrics %>%
  left_join(obs_metrics, by = c("Molecule", "Dose")) %>%
  mutate(
    FE_Cmax = Pred_Cmax / Obs_Cmax,
    FE_AUC = Pred_AUC / Obs_AUC
  ) %>%
  arrange(Molecule, Dose, Run)

ofv_rtot_tbl <- fit_summary %>%
  select(Run, Molecule, Fit_Status, OFV, Rtot_nM, kdeg_h, kint_h, Kss_nM, CL_L_h_kg, Fit_Error) %>%
  arrange(Molecule, Run)

ofv_rtot_tbl %>%
  kable(digits = 4, caption = "OFV and estimated top-down TMDD parameters by molecule and run")

metrics_tbl %>%
  mutate(
    Obs_Cmax = round(Obs_Cmax, 2),
    Pred_Cmax = round(Pred_Cmax, 2),
    FE_Cmax = map_chr(FE_Cmax, fmt_fe),
    Obs_AUC = round(Obs_AUC, 1),
    Pred_AUC = round(Pred_AUC, 1),
    FE_AUC = map_chr(FE_AUC, fmt_fe)
  ) %>%
  select(
    Run, Molecule, Dose,
    `Observed Cmax (ug/mL)` = Obs_Cmax,
    `Predicted Cmax (ug/mL)` = Pred_Cmax,
    `FE Cmax` = FE_Cmax,
    `Observed AUC0-672h (ug*h/mL)` = Obs_AUC,
    `Predicted AUC0-672h (ug*h/mL)` = Pred_AUC,
    `FE AUC` = FE_AUC
  ) %>%
  kbl(
    escape = FALSE,
    digits = 3,
    caption = "Observed and fitted Cmax/AUC with fold error by molecule and run"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    font_size = 13
  )

write.xlsx(
  list(
    OFV_Rtot = ofv_rtot_tbl,
    Metrics = metrics_tbl,
    Fixed_TMDD_Inputs = params,
    Predictions = predictions
  ),
  file.path(outdir, "TopDownTMDD_results.xlsx"),
  overwrite = TRUE
)

saveRDS(
  list(
    fit_summary = fit_summary,
    ofv_rtot_tbl = ofv_rtot_tbl,
    metrics_tbl = metrics_tbl,
    predictions = predictions
  ),
  file.path(outdir, "TopDownTMDD_Run1_Run3_processed_Monkey.rds")
)

write.csv(ofv_rtot_tbl, file.path(outdir, "TopDownTMDD_OFV_Rtot_Monkey.csv"), row.names = FALSE)
write.csv(metrics_tbl, file.path(outdir, "TopDownTMDD_metrics_Monkey.csv"), row.names = FALSE)
```

# Diagnostic plots

```{r}
#| label: diagnostic-plots
#| fig-width: 10
#| fig-height: 7

ggplot() +
  geom_line(
    data = predictions %>%
      filter(Run == "Run3_CL_kint_fixed_Rtot", TIME <= AUC_END_H) %>%
      left_join(obs %>% distinct(ID, Molecule, Dose), by = c("ID", "Molecule")) %>%
      left_join(params %>% distinct(Molecule, Target), by = "Molecule") %>%
      arrange(Target, Molecule, Dose, TIME) %>%
      mutate(
        Facet = factor(paste(Target, Molecule, sep = " | "),
                       levels = unique(paste(Target, Molecule, sep = " | ")))
      ),
    aes(TIME, Pred_ugmL, color = factor(Dose), group = interaction(Dose, ID)),
    linewidth = 0.8
  ) +
  geom_point(
    data = obs %>%
      filter(Time_h <= AUC_END_H) %>%
      left_join(params %>% distinct(Molecule, Target), by = "Molecule") %>%
      arrange(Target, Molecule, Dose, Time_h) %>%
      mutate(
        Facet = factor(paste(Target, Molecule, sep = " | "),
                       levels = unique(paste(Target, Molecule, sep = " | ")))
      ),
    aes(Time_h, Conc_ugmL, color = factor(Dose)),
    size = 1.7
  ) +
  facet_wrap(~ Facet, scales = "free_y", ncol = 3) +
  scale_x_continuous(
    limits = c(0, AUC_END_H),
    breaks = seq(0, AUC_END_H, by = 168),
    labels = function(x) x / 24
  ) +
  scale_y_log10(
    labels = label_number(drop0trailing = TRUE),
    breaks = 10^seq(-2, 4)
  ) +
  scale_color_viridis_d(name = "Dose (mg/kg)", option = "D", end = 0.88) +
  labs(
    x = "Time post dose (days)",
    y = "Plasma Concentration (ug/mL)",
    title = "Monkey top-down QSS TMDD fitted profiles",
    subtitle = "Run 3 final model: CL and kint estimated, dataset Rtot fixed | Window = 0-28 days post dose"
  ) +
  theme_bw(base_size = 11) +
  theme(
    strip.background = element_rect(fill = "grey93"),
    strip.text = element_text(face = "bold", size = 8),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

ggsave(
  file.path(figures_dir, "TopDownTMDD_Run3_Monkey_ObsPred.png"),
  last_plot(), width = 12, height = 8, dpi = 150
)
```
