---
title: "Human Predictions using Top-down QSS TMDD Fits from Monkey"
format: html
---

# Set up

```{r}
#| label: setup
#| echo: true
#| warning: false
#| message: false

rm(list = ls())

library(tidyverse)
library(data.table)
library(scales)
library(openxlsx)
library(knitr)
library(kableExtra)
library(rxode2)

datadir <- if (dir.exists("../../01_Data")) "../../01_Data" else "01_Data"
outdir <- if (dir.exists("../SimsOutputs") || dir.exists("../Scripts")) "../SimsOutputs/" else "04_TopDown/SimsOutputs/"
dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
figures_dir <- file.path(outdir, "Figures")
dir.create(figures_dir, recursive = TRUE, showWarnings = FALSE)

RUN3_RDS <- file.path(outdir, "TopDownTMDD_Run3_CL_kint_fixed_Rtot_Monkey.rds")
PROCESSED_RDS <- file.path(outdir, "TopDownTMDD_Run1_Run3_processed_Monkey.rds")

AUC_END_H <- 28 * 24
BW_MONKEY_KG <- 3.5
BW_HUMAN_KG <- 70
CL_Q_EXPONENT <- 0.85
VOLUME_EXPONENT <- 1

normalize_names <- function(x) {
  x %>%
    str_replace_all("\\.+", ".") %>%
    str_replace_all("\\s+", ".") %>%
    str_replace_all("\\[", "") %>%
    str_replace_all("\\]", "") %>%
    str_replace_all("\\(", "") %>%
    str_replace_all("\\)", "")
}

normalize_species <- function(x) {
  case_when(
    str_to_lower(x) %in% c("nhp", "monkey", "cynomolgus monkey") ~ "Monkey",
    TRUE ~ str_to_title(x)
  )
}

unit_to_hour <- function(time, unit) {
  unit_l <- str_to_lower(str_trim(unit))
  case_when(
    unit_l %in% c("h", "hr", "hrs", "hour", "hours") ~ as.numeric(time),
    unit_l %in% c("min", "mins", "minute", "minutes") ~ as.numeric(time) / 60,
    unit_l %in% c("day", "days", "d") ~ as.numeric(time) * 24,
    TRUE ~ as.numeric(time)
  )
}

conc_ugml_to_nm <- function(conc_ugml, mw_g_mol) {
  as.numeric(conc_ugml) * 1e6 / as.numeric(mw_g_mol)
}

conc_nm_to_ugml <- function(conc_nm, mw_g_mol) {
  as.numeric(conc_nm) * as.numeric(mw_g_mol) / 1e6
}

dose_mgkg_to_nmolkg <- function(dose_mgkg, mw_g_mol) {
  as.numeric(dose_mgkg) * 1e6 / as.numeric(mw_g_mol)
}

normalize_molecule <- function(x) {
  str_to_title(str_to_lower(str_squish(as.character(x))))
}

convert_rate_to_h <- function(value, unit) {
  unit_l <- str_to_lower(str_trim(unit))
  case_when(
    unit_l %in% c("1/min", "min^-1", "1/minute") ~ as.numeric(value) * 60,
    TRUE ~ as.numeric(value)
  )
}

convert_kd_to_nm <- function(value, unit) {
  unit_l <- str_to_lower(str_trim(unit))
  unit_l <- str_replace_all(unit_l, "[^[:ascii:]]", "u")
  case_when(
    unit_l %in% c("um", "umol/l", "umol/litre", "umol/liter", "umol/l") ~ as.numeric(value) * 1000,
    unit_l %in% c("nm", "nmol/l", "nmol/litre", "nmol/liter") ~ as.numeric(value),
    TRUE ~ as.numeric(value)
  )
}

trapz_auc <- function(time, conc, t_end = Inf) {
  keep <- !is.na(time) & !is.na(conc) & conc > 0 & time <= t_end
  t <- as.numeric(time[keep])
  c <- as.numeric(conc[keep])
  if (length(t) < 2) return(NA_real_)

  ord <- order(t)
  t <- t[ord]
  c <- c[ord]
  dt <- diff(t)
  c1 <- c[-length(c)]
  c2 <- c[-1]
  use_log <- c1 != c2 & c1 > 0 & c2 > 0
  sum(ifelse(use_log, dt * (c1 - c2) / log(c1 / c2), dt * (c1 + c2) / 2))
}

fmt_fe <- function(x) {
  cell_spec(
    ifelse(is.na(x), "", formatC(x, format = "f", digits = 2)),
    background = ifelse(!is.na(x) & x >= 0.5 & x <= 2, "#d4edda", "#f8d7da"),
    color = ifelse(!is.na(x) & x >= 0.5 & x <= 2, "#155724", "#721c24"),
    bold = TRUE
  )
}

parse_infusion_duration_h <- function(route) {
  route_l <- str_to_lower(coalesce(route, ""))
  duration_h <- suppressWarnings(as.numeric(
    str_match(route_l, "([0-9]+\\.?[0-9]*)\\s*h(?:r|our)?")[, 2]
  ))
  duration_min <- suppressWarnings(as.numeric(
    str_match(route_l, "([0-9]+\\.?[0-9]*)\\s*min(?:ute)?")[, 2]
  ))
  case_when(
    str_detect(route_l, "infusion") & is.finite(duration_h) & duration_h > 0 ~ duration_h,
    str_detect(route_l, "infusion") & is.finite(duration_min) & duration_min > 0 ~ duration_min / 60,
    str_detect(route_l, "infusion") ~ 1,
    TRUE ~ 0
  )
}
```

# Read human data and inputs

```{r}
#| label: read-data

obs_raw <- read.csv(
  file.path(datadir, "Human", "mAb_data_human.csv"),
  stringsAsFactors = FALSE,
  check.names = FALSE
)
names(obs_raw) <- normalize_names(names(obs_raw))

human_param_raw <- read.xlsx(file.path(datadir, "Human", "Parameters_human.xlsx"), sheet = 1)
names(human_param_raw) <- normalize_names(names(human_param_raw))

obs_points <- obs_raw %>%
  transmute(
    Study.Id = as.character(.data[["Study.Id"]]),
    Subject.Id = coalesce(
      na_if(as.character(.data[["Subject.Id"]]), "."),
      na_if(as.character(.data[["Group.Id"]]), "."),
      as.character(.data[["Study.Id"]])
    ),
    Species = normalize_species(.data[["Species"]]),
    Dose = as.numeric(.data[["Dose"]]),
    Molecule = normalize_molecule(.data[["Molecule"]]),
    MW = as.numeric(.data[["Molecular.Weight"]]),
    Time_h = unit_to_hour(.data[["Time"]], .data[["Time.unit"]]),
    Conc_ugmL = as.numeric(.data[["Measurement"]]),
    Conc_unit = .data[["Measurement.unit"]],
    Route = .data[["Route"]],
    Infusion_duration_h = parse_infusion_duration_h(.data[["Route"]])
  ) %>%
  filter(
    Species == "Human",
    !grepl("^SC$", trimws(Route), ignore.case = TRUE),
    str_to_lower(Conc_unit) %in% c("ug/ml", "ug/ml", "microgram/ml", "micrograms/ml"),
    !is.na(Time_h),
    !is.na(Conc_ugmL),
    Conc_ugmL > 0,
    Time_h <= AUC_END_H
  ) %>%
  mutate(
    Profile = paste(Molecule, Dose, Subject.Id, Study.Id, sep = "_"),
    ID = as.integer(factor(Profile)),
    DV = conc_ugml_to_nm(Conc_ugmL, MW)
  )

obs <- obs_points %>%
  group_by(Molecule, Species, Dose, MW, Time_h) %>%
  summarise(
    Conc_ugmL = median(Conc_ugmL, na.rm = TRUE),
    DV = median(DV, na.rm = TRUE),
    Infusion_duration_h = median(Infusion_duration_h, na.rm = TRUE),
    n_points = n(),
    .groups = "drop"
  ) %>%
  arrange(Molecule, Dose, Time_h) %>%
  mutate(
    Study.Id = "Median_by_molecule_dose_time",
    Subject.Id = paste(Molecule, Dose, "median", sep = "_"),
    Route = if_else(Infusion_duration_h > 0, paste0("IV ", Infusion_duration_h, "h Infusion"), "IV"),
    Profile = paste(Molecule, Dose, sep = "_"),
    ID = as.integer(factor(Profile))
  ) %>%
  select(
    Study.Id, Subject.Id, Species, Dose, Molecule, MW, Time_h,
    Conc_ugmL, Route, Infusion_duration_h, Profile, ID, DV, n_points
  )

human_target_params <- human_param_raw %>%
  transmute(
    Molecule = normalize_molecule(.data[["Drug.Name"]]),
    Target = .data[["Target.Name"]],
    Parameter = .data[["Parameter.Name.in.PK-Sim"]],
    Value = as.numeric(.data[["Value.from.Source"]]),
    Unit = .data[["Unit"]],
    Source = .data[["Source"]]
  ) %>%
  mutate(
    parameter_key = case_when(
      str_to_lower(Parameter) == "reference concentration" ~ "human_rtot_nm",
      TRUE ~ NA_character_
    ),
    value_std = case_when(
      parameter_key == "human_rtot_nm" ~ convert_kd_to_nm(Value, Unit),
      TRUE ~ NA_real_
    )
  ) %>%
  filter(!is.na(parameter_key)) %>%
  select(Molecule, Target, parameter_key, value_std, Source) %>%
  distinct() %>%
  pivot_wider(
    names_from = parameter_key,
    values_from = value_std,
    values_fn = list(value_std = mean)
  )

obs %>%
  count(Molecule, Dose, ID, Infusion_duration_h, name = "n_time_points") %>%
  arrange(Molecule, Dose) %>%
  knitr::kable(caption = "Median human IV profiles used for monkey-to-human prediction")

human_target_params %>%
  arrange(Molecule) %>%
  knitr::kable(digits = 4, caption = "Human target reference concentrations used for QSS TMDD predictions")
```

# Read monkey Run3 parameters and scale to human

```{r}
#| label: monkey-run3-human-scaling

read_run3_summary <- function() {
  if (file.exists(RUN3_RDS)) {
    run3_fits <- readRDS(RUN3_RDS)
    run3_summary <- purrr::map_dfr(run3_fits, "summary")
    if (nrow(run3_summary) > 0) return(run3_summary)
  }

  if (file.exists(PROCESSED_RDS)) {
    processed <- readRDS(PROCESSED_RDS)
    run3_summary <- processed$fit_summary %>%
      filter(Run == "Run3_CL_kint_fixed_Rtot")
    if (nrow(run3_summary) > 0) return(run3_summary)
  }

  stop(
    "Could not read Run3 monkey parameters. Expected ",
    RUN3_RDS,
    " or ",
    PROCESSED_RDS,
    call. = FALSE
  )
}

run3_fit_summary <- read_run3_summary() %>%
  filter(Run == "Run3_CL_kint_fixed_Rtot") %>%
  mutate(Molecule = normalize_molecule(Molecule)) %>%
  select(
    Run, Molecule, OFV, monkey_Rtot_nM = Rtot_nM, kdeg_h, kint_h, Kss_nM,
    monkey_CL_L_h_kg = CL_L_h_kg,
    monkey_Vc_L_kg = Vc_L_kg,
    monkey_Q_L_h_kg = Q_L_h_kg,
    monkey_Vp_L_kg = Vp_L_kg,
    Fit_Status, Fit_Error
  ) %>%
  filter(
    if_all(
      c(kdeg_h, kint_h, Kss_nM, monkey_CL_L_h_kg, monkey_Vc_L_kg, monkey_Q_L_h_kg, monkey_Vp_L_kg),
      ~ is.finite(.x)
    )
  )

bw_ratio <- BW_HUMAN_KG / BW_MONKEY_KG
cl_q_per_kg_factor <- bw_ratio^CL_Q_EXPONENT / bw_ratio
volume_per_kg_factor <- bw_ratio^VOLUME_EXPONENT / bw_ratio

prediction_parameters <- run3_fit_summary %>%
  inner_join(human_target_params %>% select(Molecule, Target, human_Rtot_nM = human_rtot_nm), by = "Molecule") %>%
  inner_join(obs %>% distinct(Molecule, MW), by = "Molecule") %>%
  mutate(
    human_CL_L_h_kg = monkey_CL_L_h_kg * cl_q_per_kg_factor,
    human_Q_L_h_kg = monkey_Q_L_h_kg * cl_q_per_kg_factor,
    human_Vc_L_kg = monkey_Vc_L_kg * volume_per_kg_factor,
    human_Vp_L_kg = monkey_Vp_L_kg * volume_per_kg_factor,
    human_CL_L_h = monkey_CL_L_h_kg * BW_MONKEY_KG * bw_ratio^CL_Q_EXPONENT,
    human_Q_L_h = monkey_Q_L_h_kg * BW_MONKEY_KG * bw_ratio^CL_Q_EXPONENT,
    human_Vc_L = monkey_Vc_L_kg * BW_MONKEY_KG * bw_ratio^VOLUME_EXPONENT,
    human_Vp_L = monkey_Vp_L_kg * BW_MONKEY_KG * bw_ratio^VOLUME_EXPONENT,
    ksyn = human_Rtot_nM * kdeg_h
  ) %>%
  arrange(Molecule)

missing_inputs <- setdiff(unique(obs$Molecule), prediction_parameters$Molecule)
if (length(missing_inputs) > 0) {
  warning(
    "Skipping human observed molecules without both Run3 monkey fits and human target reference concentrations: ",
    paste(missing_inputs, collapse = ", "),
    call. = FALSE
  )
}

prediction_parameters %>%
  select(
    Molecule, Target,
    monkey_CL_L_h_kg, human_CL_L_h_kg, human_CL_L_h,
    monkey_Q_L_h_kg, human_Q_L_h_kg, human_Q_L_h,
    monkey_Vc_L_kg, human_Vc_L_kg, human_Vc_L,
    monkey_Vp_L_kg, human_Vp_L_kg, human_Vp_L,
    monkey_Rtot_nM, human_Rtot_nM, kdeg_h, kint_h, Kss_nM
  ) %>%
  knitr::kable(
    digits = 4,
    caption = "Monkey Run3 parameters after human allometric scaling and human target reference concentration substitution"
  )
```

# QSS TMDD human simulations

The QSS model is the same two-compartment QSS TMDD structure used for the monkey fits. Monkey Run3 `CL` and `Q` are scaled with a 0.85 allometric exponent, volumes with a 1.0 exponent, and the per-kg values are converted back to the 70 kg human basis used by the model. The estimated monkey `kint` is retained, and the target baseline is replaced by the human reference concentration.

```{r}
#| label: qss-human-simulations

qss_rxode_model <- rxode2::rxode({
  Ltot <- centr / vc
  Cp <- peri / vp
  free_drug <- 0.5 * ((Ltot - total_target - kss) + sqrt((Ltot - total_target - kss)^2 + 4 * kss * Ltot + 1e-12))
  complex <- Ltot - free_drug
  free_target <- total_target - complex

  d/dt(centr) = -cl * free_drug - q * free_drug + q * Cp - kint * complex * vc
  d/dt(peri) = q * free_drug - q * Cp
  d/dt(total_target) = ksyn - kdeg * free_target - kint * complex
})

simulate_human_qss_profile <- function(dat_id, param_row) {
  dense_times <- seq(
    0,
    max(AUC_END_H, max(dat_id$Time_h, na.rm = TRUE)),
    by = 1
  )
  event_times <- sort(unique(c(
    dense_times,
    dat_id$Time_h,
    dat_id$Infusion_duration_h[1]
  )))

  dose_nmolkg <- dose_mgkg_to_nmolkg(dat_id$Dose[1], dat_id$MW[1])
  infusion_duration <- dat_id$Infusion_duration_h[1]

  event_table <- tibble(
    id = dat_id$ID[1],
    time = event_times,
    amt = NA_real_,
    evid = 0L,
    cmt = "centr",
    rate = NA_real_
  ) %>%
    add_row(
      id = dat_id$ID[1],
      time = 0,
      amt = dose_nmolkg,
      evid = 1L,
      cmt = "centr",
      rate = if_else(infusion_duration > 0, dose_nmolkg / infusion_duration, NA_real_)
    ) %>%
    arrange(time, desc(evid)) %>%
    as.data.frame()

  theta <- c(
    cl = param_row$human_CL_L_h_kg,
    vc = param_row$human_Vc_L_kg,
    q = param_row$human_Q_L_h_kg,
    vp = param_row$human_Vp_L_kg,
    r0 = param_row$human_Rtot_nM,
    kdeg = param_row$kdeg_h,
    kint = param_row$kint_h,
    kss = param_row$Kss_nM,
    ksyn = param_row$ksyn
  )

  rxode2::rxSolve(
    qss_rxode_model,
    params = theta,
    events = event_table,
    inits = c(centr = 0, peri = 0, total_target = theta[["r0"]]),
    rtol = 1e-6,
    atol = 1e-9,
    hmin = 1e-12,
    maxsteps = 100000,
    returnType = "data.frame"
  ) %>%
    as_tibble() %>%
    transmute(
      Run = "Run3_CL_kint_fixed_Rtot_human_scaled",
      ID = as.integer(dat_id$ID[1]),
      TIME = as.numeric(time),
      Pred_nM = pmax(as.numeric(Ltot), 0),
      Pred_ugmL = conc_nm_to_ugml(Pred_nM, dat_id$MW[1]),
      Molecule = dat_id$Molecule[1],
      Dose = dat_id$Dose[1],
      MW = dat_id$MW[1],
      Infusion_duration_h = infusion_duration,
      Time_type = if_else(TIME %in% dat_id$Time_h, "Observed time", "Grid time")
    )
}

obs_for_prediction <- obs %>%
  semi_join(prediction_parameters, by = "Molecule")

predictions <- map_dfr(unique(obs_for_prediction$ID), function(profile_id) {
  dat_id <- obs_for_prediction %>%
    filter(ID == profile_id) %>%
    arrange(Time_h)
  param_row <- prediction_parameters %>%
    filter(Molecule == dat_id$Molecule[1])
  simulate_human_qss_profile(dat_id, param_row)
})

predictions_at_observed_times <- predictions %>%
  filter(Time_type == "Observed time") %>%
  inner_join(
    obs_for_prediction %>% select(ID, TIME = Time_h, Observed_ugmL = Conc_ugmL, Observed_nM = DV, n_points),
    by = c("ID", "TIME")
  )

predictions %>%
  count(Molecule, Dose, Infusion_duration_h, name = "n_simulated_times") %>%
  arrange(Molecule, Dose) %>%
  knitr::kable(caption = "Human QSS TMDD simulation audit by molecule and dose")
```

# Process outputs

```{r}
#| label: process-outputs

obs_metrics <- obs_for_prediction %>%
  group_by(Molecule, Dose, ID) %>%
  summarise(
    Obs_Cmax = max(Conc_ugmL, na.rm = TRUE),
    Obs_AUC = trapz_auc(Time_h, Conc_ugmL, t_end = AUC_END_H),
    .groups = "drop"
  ) %>%
  group_by(Molecule, Dose) %>%
  summarise(
    Obs_Cmax = mean(Obs_Cmax, na.rm = TRUE),
    Obs_AUC = mean(Obs_AUC, na.rm = TRUE),
    .groups = "drop"
  )

pred_metrics <- predictions %>%
  filter(TIME <= AUC_END_H) %>%
  group_by(Run, Molecule, Dose, ID) %>%
  summarise(
    Pred_Cmax = max(Pred_ugmL, na.rm = TRUE),
    Pred_AUC = trapz_auc(TIME, Pred_ugmL, t_end = AUC_END_H),
    .groups = "drop"
  ) %>%
  group_by(Run, Molecule, Dose) %>%
  summarise(
    Pred_Cmax = mean(Pred_Cmax, na.rm = TRUE),
    Pred_AUC = mean(Pred_AUC, na.rm = TRUE),
    .groups = "drop"
  )

metrics_tbl <- pred_metrics %>%
  left_join(obs_metrics, by = c("Molecule", "Dose")) %>%
  mutate(
    FE_Cmax = Pred_Cmax / Obs_Cmax,
    FE_AUC = Pred_AUC / Obs_AUC
  ) %>%
  arrange(Molecule, Dose, Run)

parameter_tbl <- prediction_parameters %>%
  select(
    Molecule, Target, Run, OFV, Fit_Status,
    monkey_Rtot_nM, human_Rtot_nM, kdeg_h, kint_h, Kss_nM,
    monkey_CL_L_h_kg, human_CL_L_h_kg, human_CL_L_h,
    monkey_Q_L_h_kg, human_Q_L_h_kg, human_Q_L_h,
    monkey_Vc_L_kg, human_Vc_L_kg, human_Vc_L,
    monkey_Vp_L_kg, human_Vp_L_kg, human_Vp_L
  ) %>%
  arrange(Molecule)

parameter_tbl %>%
  kable(digits = 4, caption = "Scaled human prediction parameters")

metrics_tbl %>%
  mutate(
    Obs_Cmax = round(Obs_Cmax, 2),
    Pred_Cmax = round(Pred_Cmax, 2),
    FE_Cmax = map_chr(FE_Cmax, fmt_fe),
    Obs_AUC = round(Obs_AUC, 1),
    Pred_AUC = round(Pred_AUC, 1),
    FE_AUC = map_chr(FE_AUC, fmt_fe)
  ) %>%
  select(
    Run, Molecule, Dose,
    `Observed Cmax (ug/mL)` = Obs_Cmax,
    `Predicted Cmax (ug/mL)` = Pred_Cmax,
    `FE Cmax` = FE_Cmax,
    `Observed AUC0-672h (ug*h/mL)` = Obs_AUC,
    `Predicted AUC0-672h (ug*h/mL)` = Pred_AUC,
    `FE AUC` = FE_AUC
  ) %>%
  kbl(
    escape = FALSE,
    digits = 3,
    caption = "Observed human and monkey-to-human predicted Cmax/AUC with fold error"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    font_size = 13
  )

write.xlsx(
  list(
    Human_Scaled_Parameters = parameter_tbl,
    Metrics = metrics_tbl,
    Predictions = predictions,
    Predictions_Observed_Times = predictions_at_observed_times,
    Human_Target_Inputs = human_target_params
  ),
  file.path(outdir, "TopDownTMDD_to_Human_predictions.xlsx"),
  overwrite = TRUE
)

saveRDS(
  list(
    parameter_tbl = parameter_tbl,
    metrics_tbl = metrics_tbl,
    predictions = predictions,
    predictions_at_observed_times = predictions_at_observed_times,
    human_target_params = human_target_params
  ),
  file.path(outdir, "TopDownTMDD_to_Human_predictions.rds")
)

write.csv(parameter_tbl, file.path(outdir, "TopDownTMDD_to_Human_parameters.csv"), row.names = FALSE)
write.csv(metrics_tbl, file.path(outdir, "TopDownTMDD_to_Human_metrics.csv"), row.names = FALSE)
```

# Diagnostic plots

```{r}
#| label: diagnostic-plots
#| fig-width: 10
#| fig-height: 7

ggplot() +
  geom_line(
    data = predictions %>%
      filter(TIME <= AUC_END_H) %>%
      left_join(prediction_parameters %>% distinct(Molecule, Target), by = "Molecule") %>%
      arrange(Target, Molecule, Dose, TIME) %>%
      mutate(
        Facet = factor(paste(Target, Molecule, sep = " | "),
                       levels = unique(paste(Target, Molecule, sep = " | ")))
      ),
    aes(TIME, Pred_ugmL, color = factor(Dose), group = interaction(Dose, ID)),
    linewidth = 0.8
  ) +
  geom_point(
    data = obs_for_prediction %>%
      filter(Time_h <= AUC_END_H) %>%
      left_join(prediction_parameters %>% distinct(Molecule, Target), by = "Molecule") %>%
      arrange(Target, Molecule, Dose, Time_h) %>%
      mutate(
        Facet = factor(paste(Target, Molecule, sep = " | "),
                       levels = unique(paste(Target, Molecule, sep = " | ")))
      ),
    aes(Time_h, Conc_ugmL, color = factor(Dose)),
    size = 1.7
  ) +
  facet_wrap(~ Facet, scales = "free_y", ncol = 3) +
  scale_x_continuous(
    limits = c(0, AUC_END_H),
    breaks = seq(0, AUC_END_H, by = 168),
    labels = function(x) x / 24
  ) +
  scale_y_log10(
    labels = label_number(drop0trailing = TRUE),
    breaks = 10^seq(-2, 4)
  ) +
  scale_color_viridis_d(name = "Dose (mg/kg)", option = "D", end = 0.88) +
  labs(
    x = "Time post dose (days)",
    y = "Plasma Concentration (ug/mL)",
    title = "Human top-down QSS TMDD predicted profiles",
    subtitle = "Monkey Run 3 parameters scaled to human with human target concentrations"
  ) +
  theme_bw(base_size = 11) +
  theme(
    strip.background = element_rect(fill = "grey93"),
    strip.text = element_text(face = "bold", size = 8),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

ggsave(
  file.path(figures_dir, "TopDownTMDD_to_Human_ObsPred.png"),
  last_plot(), width = 12, height = 8, dpi = 150
)
```
