---
title: "Process Simulation Outputs"
author: "NAMLink"
format:
  html:
    code-fold: true
    code-summary: "Show code"
    self-contained: true
---

# Set up

```{r}
#| label: setup
#| echo: true
#| warning: false
#| message: false

rm(list = ls())

library(tidyverse)
library(data.table)   # fast I/O (fread) and fast grouped operations
library(scales)
library(knitr)
library(openxlsx)
library(kableExtra)

# Helper functions -------------------------------------------------------------
# Keep all local helpers here so this script is stand-alone and does not depend
# on 00_functions.R.

normalize_species <- function(x) {
  dplyr::case_when(
    tolower(x) == "human" ~ "Human",
    tolower(x) %in% c("nhp", "monkey", "primate") ~ "Monkey",
    tolower(x) == "rat" ~ "Rat",
    tolower(x) == "mouse" ~ "Mouse",
    tolower(x) == "dog" ~ "Dog",
    TRUE ~ stringr::str_to_title(x)
  )
}

parse_scenario <- function(path) {
  base  <- tools::file_path_sans_ext(basename(path))
  base  <- sub("_mpk$", "", base)
  parts <- strsplit(base, "_")[[1]]
  n     <- length(parts)

  list(
    Molecule = paste(parts[seq_len(n - 2)], collapse = "_"),
    Species  = parts[n - 1],
    Dose     = suppressWarnings(as.numeric(parts[n]))
  )
}

trapz_auc <- function(time, conc, t_end = Inf) {
  keep <- !is.na(time) & !is.na(conc) & conc > 0 & time <= t_end
  t <- time[keep]
  c <- conc[keep]
  if (length(t) < 2) return(NA_real_)

  dt <- diff(t)
  c1 <- c[-length(c)]
  c2 <- c[-1]

  # Use log trapezoid on strictly declining segments; linear elsewhere.
  use_log <- c1 != c2 & c1 > 0 & c2 > 0
  trap <- ifelse(
    use_log,
    dt * (c1 - c2) / log(c1 / c2),
    dt * (c1 + c2) / 2
  )

  sum(trap)
}

fmt_fe <- function(x) {
  cell_spec(
    formatC(x, format = "f", digits = 2),
    background = ifelse(!is.na(x) & x >= 0.5 & x <= 2.0, "#d4edda", "#f8d7da"),
    color      = ifelse(!is.na(x) & x >= 0.5 & x <= 2.0, "#155724", "#721c24"),
    bold       = TRUE
  )
}

render_metrics_kbl <- function(df, caption) {
  df_fmt <- df %>%
    mutate(
      `Obs Cmax (µg/mL)`   = round(Obs_Cmax, 1),
      `Pred Cmax (µg/mL)`  = round(Pred_Cmax, 1),
      `FE Cmax`            = map_chr(FE_Cmax, fmt_fe),
      `Obs AUC (µg·h/mL)`  = round(Obs_AUC,  0),
      `Pred AUC (µg·h/mL)` = round(Pred_AUC, 0),
      `FE AUC`             = map_chr(FE_AUC,  fmt_fe)
    ) %>%
    select(
      Molecule, `Dose (mg/kg)` = Dose,
      `Obs Cmax (µg/mL)`, `Pred Cmax (µg/mL)`, `FE Cmax`,
      `Obs AUC (µg·h/mL)`, `Pred AUC (µg·h/mL)`, `FE AUC`
    )

  df_fmt %>%
    kbl(
      escape  = FALSE,
      align   = c("l", "r", "r", "r", "c", "r", "r", "c"),
      caption = caption
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE,
      font_size  = 13
    ) %>%
    column_spec(5, width = "5em") %>%
    column_spec(8, width = "5em") %>%
    pack_rows(index = setNames(
      rle(df$Molecule)$lengths,
      rle(df$Molecule)$values
    ))
}

datadir  <- "../../../01_Data/"
results_dir <- "../SimsOutputs/SimulationResults/"
figures_dir <- "../SimsOutputs/Figures"

# AUC integration window: 28 days in hours
AUC_END_H <- 28 * 24   # 672 h
```

# Load data

## Observed data

```{r}
#| label: load-obs
#| message: false

obs_raw <- read.csv(
  file.path(datadir, "Monkey", "mAb_data_clean.csv"),
  stringsAsFactors = FALSE,
  check.names      = FALSE
)

# Normalise species labels to match simulation scenario naming convention,
# convert time to hours, and keep only quantifiable rows
obs <- obs_raw %>%
  mutate(
    Species   = normalize_species(Species),
    Time_h    = case_when(
      tolower(trimws(Time.unit)) %in% c("h", "hr", "hour", "hours") ~ as.numeric(Time),
      TRUE ~ as.numeric(Time) / 60
    ),
    Conc_ugmL = as.numeric(Measurement)   # already in µg/mL
  ) %>%
  filter(!is.na(Conc_ugmL), Conc_ugmL > 0) %>%
  filter(Time_h <= 672)

# ── Human observed data (kept separate — not combined with monkey) ────────
obs_human_raw <- read.csv(
  file.path(datadir, "Human", "mAb_data_human.csv"),
  stringsAsFactors = FALSE, check.names = FALSE
)

obs_human <- obs_human_raw %>%
  mutate(
    Species   = normalize_species(Species),
    Time_h    = case_when(
      tolower(trimws(Time.unit)) %in% c("h", "hr", "hour", "hours") ~ as.numeric(Time),
      TRUE ~ as.numeric(Time) / 60
    ),
    Conc_ugmL = as.numeric(Measurement)
  ) %>%
  filter(!is.na(Conc_ugmL), Conc_ugmL > 0) %>%
  # Adalimumab human data is SC route; model is IV-only — exclude from comparison
  filter(Molecule != "Adalimumab")

# MW lookup: used for unit conversion of simulation output (µmol/L → µg/mL).
# Built from both CSVs so that Golimumab (monkey-only observed) and all human
# drugs are covered.
mw_lookup <- bind_rows(
  obs_raw       %>% distinct(Molecule, MW = Molecular.Weight),
  obs_human_raw %>% distinct(Molecule, MW = Molecular.Weight)
) %>%
  mutate(MW = as.numeric(MW)) %>%
  distinct(Molecule, .keep_all = TRUE)
```

## Simulation results — latest run

```{r}
#| label: load-sims
#| message: false

# ── Auto-select the most recently modified results sub-folder ─────────────
sim_folders <- list.dirs(results_dir, recursive = FALSE, full.names = TRUE)
if (length(sim_folders) == 0) stop("No results folders found in: ", results_dir)

latest_dir <- sim_folders[which.max(file.mtime(sim_folders))]
message("Results folder: ", latest_dir)

# ── List scenario CSVs (exclude _population parameter files) ──────────────
csv_files <- list.files(latest_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[!grepl("_population\\.csv$", csv_files)]
message(length(csv_files), " scenario CSV files found")

# ── Read CSVs ─────────────────────────────────────────────────────────────
# Expected columns: IndividualId | Time [min] | {OutputPath} [µmol/l]
# Unit conversion: µmol/L × MW (g/mol) / 1000 = µg/mL

# Function defined outside the loop so it can be exported to parallel workers
read_one_csv <- function(f) {
  meta <- parse_scenario(f)

  df <- tryCatch(
    data.table::fread(f, data.table = TRUE),
    error = function(e) NULL
  )
  if (is.null(df)) return(NULL)

  conc_col <- grep("PeripheralVenousBlood.*Plasma.*\\[", names(df), value = TRUE)
  conc_col <- conc_col[grepl("\\|mAb\\|", conc_col)]
  if (length(conc_col) == 0) return(NULL)
  conc_col <- conc_col[1]

  mw_val <- mw_lookup$MW[mw_lookup$Molecule == meta$Molecule]
  if (length(mw_val) == 0 || is.na(mw_val[1])) return(NULL)

  df[, .(
    Molecule     = meta$Molecule,
    Species      = meta$Species,
    Dose         = meta$Dose,
    IndividualId = as.integer(IndividualId),
    Time_h       = as.numeric(`Time [min]`) / 60,
    Conc_ugmL    = as.numeric(.SD[[conc_col]]) * mw_val[1] / 1000
  ), .SDcols = conc_col[1]]
}

# Parallel CSV reading — uses a socket cluster (works on Windows).
# Falls back to sequential if only 1 logical core is available.
n_cores <- min(length(csv_files), max(1L, parallel::detectCores() - 1L))
message("Reading ", length(csv_files), " scenario CSV(s) with ", n_cores, " core(s)")

if (n_cores > 1) {
  cl <- parallel::makeCluster(n_cores)
  parallel::clusterExport(cl, varlist = c("parse_scenario", "mw_lookup", "read_one_csv"),
                          envir = environment())
  parallel::clusterEvalQ(cl, library(data.table))
  sim_list <- parallel::parLapply(cl, csv_files, read_one_csv)
  parallel::stopCluster(cl)
} else {
  sim_list <- lapply(csv_files, read_one_csv)
}

sim <- data.table::rbindlist(Filter(Negate(is.null), sim_list))

if (nrow(sim) == 0) {
  stop(
    "No concentration data found in any scenario CSV.\n\n",
    "Likely cause: the OutputPath in Scenarios.xlsx references the molecule\n",
    "name (e.g. 'Atezolizumab') but the .pkml model uses a different name internally.\n\n",
    "Fix: update the OutputPaths sheet to use the correct path and re-run 02_Run_Sims.qmd."
  )
}

n_scen <- sim[, .N, by = .(Molecule, Species, Dose)][, .N]
message("Loaded ", n_scen, " scenarios | ",
        uniqueN(sim$IndividualId), " individuals per scenario")
```

# Summarise for plotting

```{r}
#| label: summarise
#| message: false
#| warning: false

# ── Simulated: 5th / 50th / 95th percentile per time point ────────────────
# data.table grouped operation is 5–20× faster than dplyr on large tables
sim_summary <- sim[, .(
  Pred_p05 = quantile(Conc_ugmL, 0.05, na.rm = TRUE),
  Pred_med = quantile(Conc_ugmL, 0.50, na.rm = TRUE),
  Pred_p95 = quantile(Conc_ugmL, 0.95, na.rm = TRUE)
), by = .(Molecule, Species, Dose, Time_h)]

# ── Observed: mean ± SD per group / time point ────────────────────────────
summarise_obs <- function(df) {
  df %>%
    group_by(Molecule, Species, Dose, Time_h) %>%
    summarise(
      Obs_mean = mean(Conc_ugmL, na.rm = TRUE),
      Obs_sd   = sd(Conc_ugmL,   na.rm = TRUE),
      n        = n(),
      .groups  = "drop"
    ) %>%
    mutate(
      Obs_lo = pmax(Obs_mean - Obs_sd, 1e-4),
      Obs_hi = Obs_mean + Obs_sd
    )
}

obs_summary       <- summarise_obs(obs)
obs_human_summary <- summarise_obs(obs_human)

# Combined only for plotting (both species appear in the same figure per drug)
obs_all_summary <- bind_rows(obs_summary, obs_human_summary)
```

# PK profiles: observed vs simulated

One figure per molecule; panels = species; colour = dose.

```{r}
#| label: pk-plots
#| fig-width: 10
#| fig-height: 4.5
#| message: false

molecules <- sort(unique(c(sim_summary$Molecule, obs_all_summary$Molecule)))

pk_plots <- lapply(molecules, function(mol) {

  sim_mol <- sim_summary[Molecule == mol]   # data.table subset
  obs_mol <- filter(obs_all_summary, Molecule == mol)

  p <- ggplot() +

    # Prediction uncertainty ribbon (5th–95th percentile)
    geom_ribbon(
      data  = sim_mol,
      aes(x = Time_h, ymin = Pred_p05, ymax = Pred_p95, fill = factor(Dose)),
      alpha = 0.18
    ) +

    # Predicted median line
    geom_line(
      data      = sim_mol,
      aes(x = Time_h, y = Pred_med, colour = factor(Dose)),
      linewidth = 0.9
    ) +

    # Observed mean points
    geom_point(
      data  = obs_mol,
      aes(x = Time_h, y = Obs_mean, colour = factor(Dose)),
      shape = 16, size = 2.3
    ) +

    # Observed ± SD error bars
    geom_errorbar(
      data      = filter(obs_mol, !is.na(Obs_sd) & n > 1),
      aes(x = Time_h, ymin = Obs_lo, ymax = Obs_hi, colour = factor(Dose)),
      width     = 0.8, linewidth = 0.45
    ) +

    facet_wrap(~ Species, scales = "free_x") +
    scale_y_log10(limits = c(0.1, NA), 
      labels = label_number(drop0trailing = TRUE),
      breaks = 10^seq(-2, 4)
    ) +
    scale_colour_viridis_d(name = "Dose (mg/kg)", option = "D", end = 0.88) +
    scale_fill_viridis_d(name   = "Dose (mg/kg)", option = "D", end = 0.88) +
    labs(
      title    = mol,
      subtitle = "Lines = simulated |  Points = observed",
      x        = "Time (h)",
      y        = "Plasma Concentration (µg/mL)"
    ) +
    theme_bw(base_size = 12) +
    theme(
      strip.background = element_rect(fill = "grey93"),
      strip.text       = element_text(face = "bold"),
      legend.position  = "bottom",
      panel.grid.minor = element_blank()
    )

  ggsave(
    file.path(figures_dir, paste0("PK_ObsPred_", mol, ".png")),
    p, width = 10, height = 4.5, dpi = 150
  )
  message("Saved plot: ", mol)
  p
})

names(pk_plots) <- molecules
pk_plots
```

# PK metrics: AUC~0–28d~ and C~max~

AUC is integrated by the linear-log trapezoidal rule over 0 –
`r AUC_END_H` h (28 days). For observed subjects the integral is
truncated at their last measured time point when that falls before 672 h.
Fold error = Predicted / Observed; values within 0.5–2.0 (2-fold) are
highlighted green.

```{r}
#| label: pk-metrics
#| message: false
#| warning: false

# ── Simulated metrics: fully vectorised — no per-group R function calls ────
# Sort once, then compute lead/lag differences across all groups simultaneously.
# This replaces thousands of trapz_auc() dispatches with three data.table ops.
setorder(sim, Molecule, Species, Dose, IndividualId, Time_h)

sim <- sim %>%
  group_by(Molecule, Species, Dose, IndividualId) %>%
  mutate(
    dt_h = Time_h - lag(Time_h),
    c_lag = lag(Conc_ugmL)
  ) %>%
  ungroup()  %>%
  mutate(
    trap = case_when(
      is.na(dt_h) | dt_h <= 0 |
        is.na(c_lag) | c_lag <= 0 |
        Conc_ugmL <= 0 ~ 0,

      c_lag != Conc_ugmL ~
        dt_h * (c_lag - Conc_ugmL) / log(c_lag / Conc_ugmL),

      TRUE ~
        dt_h * (c_lag + Conc_ugmL) / 2
    ))

sim_metrics <- sim %>%
  filter(Time_h <= AUC_END_H) %>%
  group_by(Molecule, Species, Dose, IndividualId) %>%
  summarise(
    Pred_AUC  = sum(trap, na.rm = TRUE),
    Pred_Cmax = max(Conc_ugmL, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(Molecule, Species, Dose) %>%
  summarise(
    Pred_AUC  = mean(Pred_AUC, na.rm = TRUE),
    Pred_Cmax = mean(Pred_Cmax, na.rm = TRUE),
    .groups = "drop"
  )

sim <- sim %>%
  select(-dt_h, -c_lag, -trap)

# ── Observed metrics helper: per subject → group mean ────────────────────
compute_obs_metrics <- function(df) {
  df %>%
    group_by(Molecule, Species, Dose, Subject.Id) %>%
    summarise(
      Obs_AUC  = trapz_auc(Time_h, Conc_ugmL, t_end = AUC_END_H),
      Obs_Cmax = max(Conc_ugmL, na.rm = TRUE),
      .groups  = "drop"
    ) %>%
    group_by(Molecule, Species, Dose) %>%
    summarise(
      Obs_Cmax = mean(Obs_Cmax, na.rm = TRUE),
      Obs_AUC  = mean(Obs_AUC,  na.rm = TRUE),
      .groups  = "drop"
    )
}

obs_metrics       <- compute_obs_metrics(obs)
obs_human_metrics <- compute_obs_metrics(obs_human)

compute_fe <- function(obs_m, sim_m) {
  full_join(
    as.data.frame(obs_m),
    as.data.frame(sim_m),
    by = c("Molecule", "Species", "Dose")
  ) %>%
    mutate(
      FE_Cmax = round(Pred_Cmax / Obs_Cmax, 2),
      FE_AUC  = round(Pred_AUC  / Obs_AUC,  2)
    ) %>%
    arrange(Molecule, Dose)
}

# Monkey and human joined separately — no cross-species contamination
sim_metrics_monkey <- sim_metrics %>% filter(Species == "Monkey")
sim_metrics_human  <- sim_metrics %>% filter(Species == "Human")

metrics_tbl       <- compute_fe(obs_metrics,       sim_metrics_monkey)
human_metrics_tbl <- compute_fe(obs_human_metrics, sim_metrics_human)
```

```{r}
#| label: metrics-table-nhp

metrics_tbl %>%
  render_metrics_kbl(paste("Monkey PK Metrics —",
                     paste0("AUC integrated over 0–", AUC_END_H, " h (28 days).
                            FE = Predicted / Observed.")
))
```

## Human — observed vs predicted PK metrics

Fold errors are shown for scenarios where observed human data is available at a
matching dose. Rows without observed data are shown as a priori predictions only.
Note: Pembrolizumab observed data begins at day 19 (457 h) so AUC fold error is
not meaningful (partial AUC only). Adalimumab observed data is SC 40 mg (0.57 mg/kg)
vs simulation at 3 mg/kg IV — included for reference only.

```{r}
#| label: metrics-table-human

# Scenarios with BOTH observed and predicted data (exact dose match)
human_obs <- human_metrics_tbl %>% filter(!is.na(Obs_Cmax) & !is.na(Pred_Cmax))

# Scenarios with predicted data only (no matching observed dose)
human_pred_only <- human_metrics_tbl %>%
  filter(is.na(Obs_Cmax)) %>%
  select(Molecule, `Dose (mg/kg)` = Dose,
         `Pred Cmax (µg/mL)` = Pred_Cmax,
         `Pred AUC (µg·h/mL)` = Pred_AUC) %>%
  mutate(across(where(is.numeric), ~ round(.x, 1)))

if (nrow(human_obs) > 0) {
  render_metrics_kbl(
    human_obs,
    paste("Human PK Metrics (with observed data) — AUC over 0–", AUC_END_H,
          "h. FE = Predicted / Observed.")
  )
}

if (nrow(human_pred_only) > 0) {
  human_pred_only %>%
    kbl(caption = "Human Predicted-Only PK Metrics — No observed data at these doses") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE, font_size = 13) %>%
    pack_rows(index = setNames(
      rle(human_pred_only$Molecule)$lengths,
      rle(human_pred_only$Molecule)$values
    ))
}

if (nrow(human_metrics_tbl) == 0) message("No human simulation results found.")
```

## Export to Excel

```{r}
#| label: save-metrics
#| message: false
#| output: false

metrics_out <- bind_rows(metrics_tbl, human_metrics_tbl) %>%
  arrange(Species, Molecule, Dose) %>%
  rename(
    `Dose (mg/kg)`       = Dose,
    `Obs Cmax (µg/mL)`   = Obs_Cmax,
    `Pred Cmax (µg/mL)`  = Pred_Cmax,
    `Fold Error Cmax`    = FE_Cmax,
    `Obs AUC (µg·h/mL)`  = Obs_AUC,
    `Pred AUC (µg·h/mL)` = Pred_AUC,
    `Fold Error AUC`     = FE_AUC
  ) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

wb <- createWorkbook()
addWorksheet(wb, "PK_Metrics")
writeData(wb, "PK_Metrics", metrics_out)
setColWidths(wb, "PK_Metrics",
             cols   = seq_len(ncol(metrics_out)),
             widths = c(16, 10, 13, 18, 18, 15, 18, 18, 15))

out_xlsx <- file.path(figures_dir, "PK_Metrics_ObsPred.xlsx")
saveWorkbook(wb, out_xlsx, overwrite = TRUE)
message("Saved: ", out_xlsx)
```


