---
title: "Preclinical Study Simulations: Eight-mAb Virtual Monkey Analysis"
author: NAMLink
format:
  html:
    toc: true
    code-fold: false
execute:
  warning: false
  message: false
---

## Purpose and analysis settings

This test workflow:

1. generates 100 virtual cynomolgus monkeys;
2. samples six drug/target parameters for each of eight mAbs and FcRn
   expression using Latin hypercube sampling from log-normal distributions;
3. runs 0.3, 3, and 10 mg/kg simulations for every mAb and monkey;
4. calculates both rich-grid ("true") and experimentally sparse AUC~0-672h~;
   and
5. estimates power to detect a two-fold observed-versus-PBPK-predicted AUC
   difference.

The rich-grid PBPK AUC is treated as the predicted reference distribution.
Simulated "observed" AUCs are estimated from sparse profiles, include
bioanalytical/sampling error, and have a geometric mean two-fold above the
PBPK prediction under the power-analysis alternative.

```{r}
#| label: setup

rm(list = ls())

suppressPackageStartupMessages({
  library(tidyverse)
  library(openxlsx)
  library(lhs)
  library(ospsuite)
  library(esqlabsR)
  library(knitr)
  library(kableExtra)
})

set.seed(20260624)

N_VIRTUAL       <- 100L
PARAMETER_CV    <- 0.30
FCRN_CV         <- 0.30
OBSERVATION_CV  <- 0.15
EFFECT_FOLD     <- 2
AUC_END_H       <- 28 * 24
ALPHA           <- 0.05
N_POWER_SIM     <- 1000L  # increase to >=5000 for the final analysis
N_SPARSE_REP    <- 100L
RUN_PBPK        <- TRUE # set TRUE only when PBPK simulations must be regenerated

STUDY_DOSES     <- c(0.3, 3, 10)
N_PER_GROUP     <- 2:5
FIXED_TOTAL_N   <- 4:15

MABS <- c(
  "Adalimumab", "Atezolizumab", "Avelumab", "Bevacizumab",
  "Golimumab", "Nivolumab", "Pembrolizumab", "Tocilizumab"
)

# Common plotting style. Figures are intentionally compact, while retaining a
# base font size of 12 so labels remain readable in the rendered report.
STUDY_THEME <- theme_bw(base_size = 12) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

# Reduced serial sampling schedule selected for this study-design exercise.
# It retains an early distribution-phase sample at 15 min, two additional
# samples within the first 6 h, and six samples spanning Days 2-28.
#
# The schedule is sparser than the published Gupta et al. cynomolgus mAb
# schedule (15 min, 4 h, 8 h; Days 1, 3, 5, 7, 10, 14, 21, 28 and 42).
# Consequently, the present analysis explicitly evaluates its bias and
# variability against the rich PBPK AUC rather than claiming that this exact
# reduced schedule is a literature-standard schedule.
#
# Reference:
# Gupta P et al. MAbs. 2016;8(5):991-997.
# doi:10.1080/19420862.2016.1167294; PMID:27031797.
SPARSE_TIME_H <- c(
  0.25, 4, 6,
  2, 6, 10, 14, 21, 28
) * c(1, 1, 1, rep(24, 6))

study_dir       <- normalizePath(".", winslash = "/", mustWork = TRUE)
project_dir     <- normalizePath("..", winslash = "/", mustWork = TRUE)
data_dir        <- file.path(project_dir, "01_Data", "Monkey")
config_dir      <- file.path(study_dir, "Configurations")
population_dir  <- file.path(config_dir, "PopulationsCSV")
output_root     <- file.path(study_dir, "SimsOutputs")
figure_dir      <- file.path(output_root, "Figures")
table_dir       <- file.path(output_root, "Tables")
simulation_results_dir <- file.path(output_root, "SimulationResults")

walk(
  c(
    config_dir, population_dir, output_root, figure_dir, table_dir,
    simulation_results_dir
  ),
  ~ dir.create(.x, recursive = TRUE, showWarnings = FALSE)
)

safe_unlink_within <- function(paths, root, recursive = FALSE) {
  paths <- paths[file.exists(paths)]
  if (length(paths) == 0L) return(invisible(character()))

  normalized_root <- normalizePath(root, winslash = "/", mustWork = TRUE)
  normalized_paths <- normalizePath(paths, winslash = "/", mustWork = TRUE)
  inside_root <- startsWith(
    paste0(normalized_paths, "/"),
    paste0(normalized_root, "/")
  )
  if (!all(inside_root)) {
    stop("Refusing to delete path outside ", normalized_root)
  }

  unlink(normalized_paths, recursive = recursive, force = TRUE)
  if (any(file.exists(normalized_paths))) {
    stop(
      "Could not delete: ",
      paste(normalized_paths[file.exists(normalized_paths)], collapse = ", ")
    )
  }
  invisible(normalized_paths)
}

# Remove figures no longer produced by this workflow so stale files cannot be
# mistaken for outputs from the current run.
safe_unlink_within(
  file.path(
    figure_dir,
    c(
      "All_mAbs_sparse_vs_rich_AUC.png",
      "All_mAbs_power_vs_sample_size.png"
    )
  ),
  output_root
)

safe_save_workbook <- function(wb, path) {
  tmp <- tempfile(fileext = ".xlsx", tmpdir = config_dir)
  on.exit(unlink(tmp), add = TRUE)
  saveWorkbook(wb, tmp, overwrite = TRUE)
  if (!file.copy(tmp, path, overwrite = TRUE)) {
    stop("Could not write ", path, ". Close the workbook if it is open in Excel.")
  }
  invisible(path)
}

make_lognormal <- function(u, typical, cv) {
  # The spreadsheet value is interpreted as the typical value (median).
  sdlog <- sqrt(log1p(cv^2))
  qlnorm(u, meanlog = log(typical), sdlog = sdlog)
}

trapz_auc <- function(time_h, concentration, end_h = Inf) {
  keep <- is.finite(time_h) & is.finite(concentration) &
    time_h <= end_h & concentration >= 0
  time_h <- time_h[keep]
  concentration <- concentration[keep]
  ord <- order(time_h)
  time_h <- time_h[ord]
  concentration <- concentration[ord]
  if (length(time_h) < 2L) return(NA_real_)
  sum(diff(time_h) * (head(concentration, -1) + tail(concentration, -1)) / 2)
}

linup_logdown_auc <- function(time_h, concentration) {
  # Standard linear-up/log-down trapezoidal NCA:
  # - linear trapezoid when concentrations rise or either value is nonpositive;
  # - logarithmic trapezoid when concentrations decline.
  keep <- is.finite(time_h) & is.finite(concentration) & concentration >= 0
  time_h <- time_h[keep]
  concentration <- concentration[keep]
  ord <- order(time_h)
  time_h <- time_h[ord]
  concentration <- concentration[ord]
  if (length(time_h) < 2L) return(NA_real_)

  dt <- diff(time_h)
  c1 <- head(concentration, -1)
  c2 <- tail(concentration, -1)
  declining <- c2 < c1 & c1 > 0 & c2 > 0
  intervals <- dt * (c1 + c2) / 2
  intervals[declining] <-
    dt[declining] * (c1[declining] - c2[declining]) /
    log(c1[declining] / c2[declining])
  sum(intervals)
}
```

## Generate the monkey virtual population

The CSV is written in the format consumed by `ospsuite::loadPopulation()` and
`esqlabsR` when `ReadPopulationFromCSV = TRUE`. The simulations below use the
same generated monkeys as individual scenarios so that each monkey can receive
its own parameter sheet from `Individuals.xlsx`.

```{r}
#| label: generate-population

population_characteristics <- createPopulationCharacteristics(
  species             = Species$Monkey,
  population          = "Monkey",
  numberOfIndividuals = N_VIRTUAL,
  proportionOfFemales = 50,
  weightMin           = 2,
  weightMax           = 8,
  weightUnit          = "kg",
  ageMin              = 2,
  ageMax              = 30,
  ageUnit             = "year(s)",
  seed                 = 20260624L
)

monkey_population <- createPopulation(population_characteristics)
monkey_population_df <- populationToDataFrame(monkey_population$population)
weight_column <- grep(
  "^Organism\\|Weight",
  names(monkey_population_df),
  value = TRUE
)
stopifnot(length(weight_column) == 1L)

population_csv <- file.path(
  population_dir, paste0("Monkey-Base-n", N_VIRTUAL, ".csv")
)
write.csv(
  monkey_population_df,
  population_csv,
  row.names = FALSE,
  na = ""
)

# Confirm that the saved population can be imported by ospsuite.
population_import_check <- loadPopulation(population_csv)
stopifnot(length(population_import_check$allIndividualIds) == N_VIRTUAL)

monkey_population_df %>%
  transmute(
    IndividualId,
    Gender,
    Weight_kg = .data[[weight_column]]
  ) %>%
  kable(caption = "Generated 100-monkey base virtual population") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = FALSE)
```

## Sample mAb parameters and shared physiological IIV by LHS

All six positive drug/target parameters in
`01_Data/Monkey/Parameters.xlsx` are sampled. We additionally sample the
physiological FcRn abundance parameter
`Organism|EndogenousIgG|Endosome|Start concentration of free FcRn (endosome)`.

The AntiPD1 monkey model uses a typical value of 21 µmol/L. FcRn expression
controls receptor abundance available for IgG salvage, whereas the existing
`Kd (FcRn)` controls each mAb's affinity for that receptor. Both
are sampled independently with 40% CV. Independence is an exploratory
assumption because no covariance information is available.

```{r}
#| label: sample-parameters

drug_parameter_source <- read.xlsx(
  file.path(data_dir, "Parameters.xlsx"),
  sheet = "Included",
  check.names = FALSE
) %>%
  rename(
    Drug = Drug.Name,
    Target = Target.Name,
    Parameter = `Parameter.Name.in.PK-Sim`,
    Typical = Value.from.Source
  ) %>%
  filter(Drug %in% MABS) %>%
  mutate(
    Typical = as.numeric(Typical),
    CV = PARAMETER_CV,
    Parameter_class = "Drug/target"
  )

stopifnot(
  nrow(drug_parameter_source) == 6L * length(MABS),
  all(drug_parameter_source$Typical > 0)
)

# FcRn expression is an organism-level physiological parameter traced directly
# in the monkey AntiPD1 PKML. A +30% perturbation changed AUC by about +9%,
# confirming that this path is active in the antibody PBPK model.
fcrn_source <- tibble(
  Drug = "Physiology",
  Target = "FcRn",
  Parameter = "Start concentration of free FcRn (endosome)",
  Typical = 21,
  CV = FCRN_CV,
  Parameter_class = "Physiology"
)

container_map <- c(
  "Kd (FcRn) in endosomal space" = "mAb",
  "Kd" = "mAb-Target-InVitro",
  "koff" = "mAb-Target-InVitro",
  "kint" = "Complex Internalization",
  "kdeg" = "Target degradation",
  "Reference concentration" = "Target"
)

unit_map <- c(
  "Kd (FcRn) in endosomal space" = "µmol/l",
  "Kd" = "µmol/l",
  "koff" = "1/min",
  "kint" = "1/h",
  "kdeg" = "1/h",
  "Reference concentration" = "nmol/l"
)

container_map <- c(
  container_map,
  "Start concentration of free FcRn (endosome)" =
    "Organism|EndogenousIgG|Endosome"
)
unit_map <- c(
  unit_map,
  "Start concentration of free FcRn (endosome)" = "µmol/l"
)

# Physiological draws are shared across mAbs for a given virtual monkey.
physiology_source <- fcrn_source
physiology_lhs <- randomLHS(N_VIRTUAL, nrow(physiology_source))
physiology_draws <- map2_dfc(
  seq_len(nrow(physiology_source)),
  seq_len(nrow(physiology_source)),
  ~ make_lognormal(
    physiology_lhs[, .x],
    typical = physiology_source$Typical[.y],
    cv = physiology_source$CV[.y]
  )
)
names(physiology_draws) <- physiology_source$Parameter
physiology_draws <- physiology_draws %>%
  mutate(IndividualId = seq_len(N_VIRTUAL) - 1L, .before = 1)

# Drug/target parameters are sampled independently for each mAb.
drug_parameter_draws <- map_dfr(MABS, function(drug) {
  source_rows <- drug_parameter_source %>%
    filter(Drug == drug)
  lhs_matrix <- randomLHS(N_VIRTUAL, nrow(source_rows))
  draws <- map2_dfc(
    seq_len(nrow(source_rows)),
    seq_len(nrow(source_rows)),
    ~ make_lognormal(
      lhs_matrix[, .x],
      typical = source_rows$Typical[.y],
      cv = source_rows$CV[.y]
    )
  )
  names(draws) <- source_rows$Parameter
  draws %>%
    mutate(
      Drug = drug,
      IndividualId = seq_len(N_VIRTUAL) - 1L,
      .before = 1
    ) %>%
    pivot_longer(
      -c(Drug, IndividualId),
      names_to = "Parameter",
      values_to = "Value"
    ) %>%
    left_join(
      source_rows %>%
        select(Parameter, Typical, CV, Parameter_class),
      by = "Parameter"
    )
})

sampling_qc <- bind_rows(
  drug_parameter_draws,
  physiology_draws %>%
    pivot_longer(
      -IndividualId, names_to = "Parameter", values_to = "Value"
    ) %>%
    left_join(physiology_source, by = "Parameter") %>%
    mutate(Drug = "Shared physiology")
) %>%
  group_by(Drug, Parameter) %>%
  summarise(
    Typical = first(Typical),
    Intended_CV_percent = 100 * first(CV),
    Sample_median = median(Value),
    Sample_CV_percent = 100 * sd(Value) / mean(Value),
    Parameter_class = first(Parameter_class),
    .groups = "drop"
  )

sampling_qc %>%
  kable(digits = 4, caption = "LHS parameter-sampling check") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = FALSE)
```

## Write esqlabsR configuration files

At this scale, individual parameter draws are stored directly in one population
CSV per mAb. The same 100 physiological monkeys are reused across mAbs, while
drug/target parameters are sampled separately for each molecule. Within an mAb,
the same individual draw is reused at all three doses.

```{r}
#| label: write-configurations

# At n=100 across eight mAbs, individual-scenario workbooks would require 800
# parameter sheets and 2,400 separate simulations. Instead, each mAb receives
# one PK-Sim population CSV containing 100 individuals and their parameter
# columns. esqlabsR then runs only 24 population scenarios (8 mAbs × 3 doses)
# while retaining all individual-level outputs.
population_file_index <- map_dfr(MABS, function(drug) {
  drug_wide <- drug_parameter_draws %>%
    filter(Drug == drug) %>%
    select(IndividualId, Parameter, Value) %>%
    pivot_wider(names_from = Parameter, values_from = Value)

  pop <- monkey_population_df %>%
    left_join(drug_wide, by = "IndividualId") %>%
    left_join(physiology_draws, by = "IndividualId")

  # Rename sampled drug parameters to complete PK-Sim parameter paths.
  for (parameter_name in unique(
    drug_parameter_source$Parameter[drug_parameter_source$Drug == drug]
  )) {
    full_path <- paste(
      container_map[[parameter_name]], parameter_name, sep = "|"
    )
    pop[[full_path]] <- pop[[parameter_name]]
    pop[[parameter_name]] <- NULL
  }

  pop[[
    "Organism|EndogenousIgG|Endosome|Start concentration of free FcRn (endosome)"
  ]] <- pop[["Start concentration of free FcRn (endosome)"]]
  pop[["Start concentration of free FcRn (endosome)"]] <- NULL

  population_id <- paste0(drug, "_Monkey_n", N_VIRTUAL)
  population_path <- file.path(population_dir, paste0(population_id, ".csv"))
  write.csv(pop, population_path, row.names = FALSE, na = "")

  # Round-trip validation catches malformed path columns before simulation.
  imported_population <- loadPopulation(population_path)
  stopifnot(length(imported_population$allIndividualIds) == N_VIRTUAL)

  tibble(
    Drug = drug,
    PopulationId = population_id,
    PopulationFile = population_path
  )
})

# Individuals.xlsx remains part of the standard project configuration but is
# not used by population scenarios.
individual_wb <- createWorkbook()
addWorksheet(individual_wb, "IndividualBiometrics")
writeData(
  individual_wb, "IndividualBiometrics",
  data.frame(
    IndividualId = character(), Species = character(),
    Population = character(), Gender = character(),
    `Weight [kg]` = numeric(), `Height [cm]` = numeric(),
    `Age [year(s)]` = numeric(), `Protein Ontogenies` = character(),
    check.names = FALSE
  )
)
safe_save_workbook(individual_wb, file.path(config_dir, "Individuals.xlsx"))

# Empty Global sheet: drug/target parameters and all physiological IIV
# parameters are stored on the individual-specific sheets.
model_wb <- createWorkbook()
addWorksheet(model_wb, "Global")
writeData(
  model_wb, "Global",
  data.frame(
    `Container Path` = character(),
    `Parameter Name` = character(),
    Value = numeric(),
    Units = character(),
    check.names = FALSE
  )
)
safe_save_workbook(model_wb, file.path(config_dir, "ModelParameters.xlsx"))

population_config <- data.frame(
  PopulationName = paste0("Monkey-Base-n", N_VIRTUAL),
  species = "Monkey",
  population = "Monkey",
  numberOfIndividuals = N_VIRTUAL,
  proportionOfFemales = 50,
  weightMin = 2,
  weightMax = 8,
  weightUnit = "kg",
  heightMin = NA_real_,
  heightMax = NA_real_,
  heightUnit = NA_character_,
  ageMin = 2,
  ageMax = 30,
  BMIMin = NA_real_,
  BMIMax = NA_real_,
  BMIUnit = NA_character_,
  `Protein Ontogenies` = NA_character_,
  check.names = FALSE
)
user_variability <- data.frame(
  `Container Path` = character(),
  `Parameter Name` = character(),
  Mean = numeric(),
  SD = numeric(),
  Distribution = character(),
  check.names = FALSE
)
write.xlsx(
  list(
    Demographics = population_config,
    UserDefinedVariability = user_variability
  ),
  file.path(config_dir, "Populations.xlsx"),
  overwrite = TRUE
)

application_wb <- createWorkbook()
for (drug in MABS) {
for (dose in STUDY_DOSES) {
  protocol <- paste0(drug, "_", dose, "_mpk")
  addWorksheet(application_wb, protocol)
  writeData(
    application_wb, protocol,
    data.frame(
      `Container Path` =
        "Events|IV_Bolus|Application_1|ProtocolSchemaItem",
      `Parameter Name` = "DosePerBodyWeight",
      Value = dose,
      Units = "mg/kg",
      `Infusion time` = 1,
      `Infusion time unit` = "min",
      check.names = FALSE
    )
  )
}
}
safe_save_workbook(application_wb, file.path(config_dir, "Applications.xlsx"))

scenario_grid <- crossing(
  Drug = MABS,
  Dose_mgkg = STUDY_DOSES
) %>%
  left_join(population_file_index, by = "Drug") %>%
  left_join(
    drug_parameter_source %>%
      distinct(Drug, Target),
    by = "Drug"
  ) %>%
  mutate(
    Scenario_name = paste(Drug, "Monkey", paste0("Dose", Dose_mgkg), sep = "_"),
    IndividualId = NA_character_,
    ReadPopulationFromCSV = TRUE,
    ModelParameterSheets = '"Global"',
    ApplicationProtocol = paste0(Drug, "_", Dose_mgkg, "_mpk"),
    SimulationTime = "0, 12, 12; 13, 672, 28",
    SimulationTimeUnit = "h",
    SteadyState = FALSE,
    SteadyStateTime = NA_real_,
    SteadyStateTimeUnit = NA_character_,
    ModelFile = case_when(
      Target == "PD-L1" ~
        "PKML_Files_Monkey/PKML_Files_Monkey/AntiPDL1.pkml",
      Target == "PD-1" ~
        "PKML_Files_Monkey/PKML_Files_Monkey/AntiPD1.pkml",
      Target == "VEGFA" ~
        "PKML_Files_Monkey/PKML_Files_Monkey/AntiVEGFA.pkml",
      Target == "IL-6R" ~
        "PKML_Files_Monkey/PKML_Files_Monkey/AntiIL6R.pkml",
      Target == "TNFA" ~
        "PKML_Files_Monkey/PKML_Files_Monkey/AntiTNFA.pkml"
    ),
    OutputPathsIds = "PVB"
  ) %>%
  select(
    Scenario_name, IndividualId, PopulationId, ReadPopulationFromCSV,
    ModelParameterSheets, ApplicationProtocol, SimulationTime,
    SimulationTimeUnit, SteadyState, SteadyStateTime, SteadyStateTimeUnit,
    ModelFile, OutputPathsIds
  )

output_paths <- data.frame(
  OutputPathId = "PVB",
  OutputPath =
    "Organism|PeripheralVenousBlood|mAb|Plasma (Peripheral Venous Blood)",
  stringsAsFactors = FALSE
)
write.xlsx(
  list(Scenarios = scenario_grid, OutputPaths = output_paths),
  file.path(config_dir, "Scenarios.xlsx"),
  overwrite = TRUE
)

project_configuration <- data.frame(
  Property = c(
    "modelFolder", "configurationsFolder", "modelParamsFile",
    "individualsFile", "populationsFile", "populationsFolder",
    "scenariosFile", "applicationsFile", "plotsFile", "dataFolder",
    "dataFile", "dataImporterConfigurationFile", "outputFolder"
  ),
  Value = c(
    "../00_Models/", "Configurations/", "ModelParameters.xlsx",
    "Individuals.xlsx", "Populations.xlsx", "PopulationsCSV",
    "Scenarios.xlsx", "Applications.xlsx", NA, NA, NA, NA, "SimsOutputs/"
  ),
  Description = NA_character_
)
write.xlsx(
  project_configuration,
  file.path(study_dir, "ProjectConfiguration.xlsx"),
  overwrite = TRUE
)

configuration_qc <- tibble(
  File = c(
    "ProjectConfiguration.xlsx", "Configurations/Individuals.xlsx",
    "Configurations/ModelParameters.xlsx", "Configurations/Populations.xlsx",
    "Configurations/Applications.xlsx", "Configurations/Scenarios.xlsx",
    paste0(
      "Configurations/PopulationsCSV/",
      population_file_index$PopulationId, ".csv"
    )
  ),
  Exists = file.exists(file.path(study_dir, File))
)
stopifnot(
  all(configuration_qc$Exists),
  nrow(scenario_grid) == length(MABS) * length(STUDY_DOSES)
)
configuration_qc %>% kable() %>% kable_styling(full_width = FALSE)
```

## Run all eight mAb population simulations

Set `RUN_PBPK <- FALSE` in the setup chunk to reuse the newest completed result
folder without rerunning PK-Sim.

```{r}
#| label: run-pbpk

project <- createProjectConfiguration(
  file.path(study_dir, "ProjectConfiguration.xlsx")
)

if (RUN_PBPK) {
  # A fresh PBPK run must never coexist with prior scenario-result folders.
  prior_result_entries <- list.files(
    simulation_results_dir,
    full.names = TRUE,
    all.files = TRUE,
    no.. = TRUE
  )
  safe_unlink_within(
    prior_result_entries,
    simulation_results_dir,
    recursive = TRUE
  )

  scenario_configurations <- readScenarioConfigurationFromExcel(
    scenarioNames = scenario_grid$Scenario_name,
    projectConfiguration = project
  )
  scenarios <- createScenarios(scenario_configurations)

  run_options <- SimulationRunOptions$new()
  run_options$checkForNegativeValues <- FALSE
  run_options$numberOfCores <- min(
    8L, max(1L, parallel::detectCores(logical = TRUE) - 1L)
  )

  simulation_results <- runScenarios(
    scenarios = scenarios,
    simulationRunOptions = run_options
  )
  result_folder <- saveScenarioResults(simulation_results, project)

  # esqlabsR may copy population CSVs and individual PKML files beneath the
  # output folder while running. They are inputs/intermediates, not retained
  # study results. Preserve the scenario result CSVs in result_folder.
  output_pkml_files <- list.files(
    output_root,
    pattern = "\\.pkml$",
    recursive = TRUE,
    full.names = TRUE,
    ignore.case = TRUE
  )
  population_csv_names <- paste0(
    population_file_index$PopulationId, ".csv"
  )
  output_population_csvs <- list.files(
    output_root,
    pattern = "\\.csv$",
    recursive = TRUE,
    full.names = TRUE,
    ignore.case = TRUE
  )
  output_population_csvs <- output_population_csvs[
    basename(output_population_csvs) %in% population_csv_names |
      str_detect(
        str_replace_all(output_population_csvs, "\\\\", "/"),
        "/PopulationsCSV/"
      )
  ]
  safe_unlink_within(
    c(output_pkml_files, output_population_csvs),
    output_root
  )
} else {
  candidate_folders <- list.dirs(
    file.path(output_root, "SimulationResults"),
    recursive = FALSE,
    full.names = TRUE
  )
  if (length(candidate_folders) == 0L) {
    stop("No prior simulation result folder found. Set RUN_PBPK <- TRUE.")
  }
  result_folder <- candidate_folders[which.max(file.info(candidate_folders)$mtime)]
}

result_folder
```

## Rich-grid truth and reduced sparse-sampling AUC

The model output grid is retained to calculate a numerical reference or
"true PBPK AUC." It is not treated as an experimentally achievable sampling
schedule.

The experimental analysis instead samples each PBPK profile at the reduced
schedule selected for this exercise:

`0.25, 4, 6 h; Days 2, 6, 10, 14, 21 and 28`

This schedule is intentionally sparser than the single-IV-dose cynomolgus
monkey mAb schedule reported by Gupta et al. The literature schedule provides
the biological context, while the exact reduced schedule above is a study
design choice being evaluated against the rich PBPK reference.

At each scheduled time, the PBPK concentration is obtained by linear
interpolation. A mean-one log-normal observation multiplier with 40% CV is
then added. This represents combined blood-sampling and ligand-binding-assay
variability; it is separate from biological IIV in the PBPK parameters.
Wang et al. demonstrated that serial microsampling can recover cynomolgus mAb
PK parameters comparable to conventional serum sampling, supporting the use
of serial sparse profiles in this setting.

Because the first post-dose sample is at 15 minutes, concentration from time
zero to 15 minutes is approximated by the 15-minute concentration. For a
long-half-life mAb this interval contributes minimally to AUC~0-28d~, but the
assumption is stated explicitly. Sparse AUC is calculated with the standard
linear-up/log-down trapezoidal method.

References:

- Gupta P, et al. *MAbs*. 2016;8:991-997.
  doi:10.1080/19420862.2016.1167294; PMID:27031797.
- Wang Y, et al. *Pharm Res*. 2021;38:819-830.
  doi:10.1007/s11095-021-03044-6; PMID:33982224.

```{r}
#| label: auc-extraction

result_files <- list.files(
  result_folder,
  pattern = "_Monkey_Dose.*\\.csv$",
  full.names = TRUE
)
result_files <- result_files[
  !str_detect(result_files, "_population\\.csv$")
]
stopifnot(length(result_files) == nrow(scenario_grid))

processed_profiles <- map(result_files, function(path) {
  scenario_name <- tools::file_path_sans_ext(basename(path))
  drug <- stringr::str_match(
    scenario_name,
    "^(.+)_Monkey_Dose"
  )[, 2]
  dose_mgkg <- as.numeric(stringr::str_match(
    scenario_name,
    "_Dose([0-9.]+)$"
  )[, 2])
  stopifnot(is.finite(dose_mgkg), drug %in% MABS)
  # Population result files are large (~50 MB each). Process one scenario at
  # a time with data.table and retain only individual AUC plus nine scheduled
  # concentrations, avoiding a >40-million-row combined rich-profile object.
  result <- data.table::fread(path, check.names = FALSE)
  concentration_column <- grep(
    "Plasma \\(Peripheral Venous Blood\\)",
    names(result),
    value = TRUE
  )
  stopifnot(length(concentration_column) == 1L)

  data.table::setnames(
    result,
    c("Time [min]", concentration_column),
    c("Time_min", "Concentration_umol_L")
  )
  result <- result[Time_min <= AUC_END_H * 60]

  rich <- result[, .(
    Rich_AUC_umol_h_L = trapz_auc(
      Time_min / 60, Concentration_umol_L
    )
  ), by = IndividualId] %>%
    as_tibble() %>%
    mutate(
      Scenario_name = scenario_name,
      Drug = drug,
      Dose_mgkg = dose_mgkg,
      .before = 1
    )

  sparse <- result[, {
    interpolated <- approx(
      x = Time_min / 60,
      y = Concentration_umol_L,
      xout = SPARSE_TIME_H,
      rule = 2
    )$y
    .(Time_h = SPARSE_TIME_H,
      True_concentration_umol_L = interpolated)
  }, by = IndividualId] %>%
    as_tibble() %>%
    mutate(
      Scenario_name = scenario_name,
      Drug = drug,
      Dose_mgkg = dose_mgkg,
      .before = 1
    )

  list(rich = rich, sparse = sparse)
})

rich_auc <- map_dfr(processed_profiles, "rich")
sparse_concentrations_true <- map_dfr(processed_profiles, "sparse")
rm(processed_profiles)
invisible(gc())

# Generate repeated experimental realizations from each biological monkey.
# rlnorm(meanlog = -sigma^2/2) gives a multiplier with arithmetic mean 1.
# Thus observation error adds variance without systematically inflating
# concentration or AUC.
sigma_observation <- sqrt(log1p(OBSERVATION_CV^2))

sparse_auc_replicates <- crossing(
  sparse_concentrations_true %>%
    distinct(Scenario_name, Drug, IndividualId, Dose_mgkg),
  SparseReplicate = seq_len(N_SPARSE_REP)
) %>%
  left_join(
    sparse_concentrations_true,
    by = c("Scenario_name", "Drug", "IndividualId", "Dose_mgkg")
  ) %>%
  mutate(
    Observed_concentration_umol_L =
      True_concentration_umol_L *
      rlnorm(
        n(),
        meanlog = -0.5 * sigma_observation^2,
        sdlog = sigma_observation
      )
  ) %>%
  group_by(
    Scenario_name, Drug, IndividualId, Dose_mgkg, SparseReplicate
  ) %>%
  summarise(
    # Flat back-extrapolation over the unobserved 0-0.25 h interval.
    Sparse_AUC_umol_h_L = linup_logdown_auc(
      c(0, Time_h),
      c(first(Observed_concentration_umol_L),
        Observed_concentration_umol_L)
    ),
    .groups = "drop"
  ) %>%
  left_join(
    rich_auc,
    by = c("Scenario_name", "Drug", "IndividualId", "Dose_mgkg")
  ) %>%
  mutate(
    Sparse_to_rich_ratio = Sparse_AUC_umol_h_L / Rich_AUC_umol_h_L,
    Sparse_bias_percent = 100 * (Sparse_to_rich_ratio - 1)
  )

stopifnot(
  nrow(rich_auc) == length(MABS) * length(STUDY_DOSES) * N_VIRTUAL,
  nrow(sparse_auc_replicates) ==
    length(MABS) * length(STUDY_DOSES) * N_VIRTUAL * N_SPARSE_REP,
  all(is.finite(sparse_auc_replicates$Sparse_AUC_umol_h_L)),
  all(sparse_auc_replicates$Sparse_AUC_umol_h_L > 0)
)

write.csv(
  sparse_auc_replicates,
  file.path(table_dir, "All_mAbs_sparse_AUC_replicates.csv"),
  row.names = FALSE
)

# This empirical sparse-AUC distribution is used in all subsequent power
# analyses. It combines biological variability among virtual monkeys with
# uncertainty from sparse sampling and bioanalysis.
auc_individual <- sparse_auc_replicates %>%
  transmute(
    Scenario_name,
    Drug,
    IndividualId,
    Dose_mgkg,
    SparseReplicate,
    AUC_umol_h_L = Sparse_AUC_umol_h_L
  )

# Cache each drug-dose reference vector once. This avoids repeatedly filtering
# the 240,000-row sparse-AUC table inside tens of thousands of Monte Carlo
# replicates.
reference_key <- function(drug, dose) paste(drug, dose, sep = "__")
rich_reference_list <- split(
  rich_auc$Rich_AUC_umol_h_L,
  reference_key(rich_auc$Drug, rich_auc$Dose_mgkg)
)
sparse_reference_list <- split(
  auc_individual$AUC_umol_h_L,
  reference_key(auc_individual$Drug, auc_individual$Dose_mgkg)
)

auc_summary <- sparse_auc_replicates %>%
  group_by(Drug, Dose_mgkg) %>%
  summarise(
    Biological_monkeys = n_distinct(IndividualId),
    Sparse_profiles = n(),
    Rich_AUC_geometric_mean =
      exp(mean(log(Rich_AUC_umol_h_L))),
    Rich_AUC_CV_percent =
      100 * sd(
        Rich_AUC_umol_h_L[!duplicated(
          paste(Scenario_name, IndividualId)
        )]
      ) / mean(
        Rich_AUC_umol_h_L[!duplicated(
          paste(Scenario_name, IndividualId)
        )]
      ),
    Sparse_AUC_geometric_mean =
      exp(mean(log(Sparse_AUC_umol_h_L))),
    Sparse_AUC_CV_percent =
      100 * sd(Sparse_AUC_umol_h_L) / mean(Sparse_AUC_umol_h_L),
    Median_sparse_bias_percent = median(Sparse_bias_percent),
    Sparse_bias_5th_percentile = quantile(Sparse_bias_percent, 0.05),
    Sparse_bias_95th_percentile = quantile(Sparse_bias_percent, 0.95),
    .groups = "drop"
  )

auc_summary %>%
  kable(
    digits = 3,
    caption = paste0(
      "Sparse versus rich PBPK AUC0-672h. Sparse profiles use ",
      scales::percent(OBSERVATION_CV),
      " proportional observation CV."
    )
  ) %>%
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = FALSE)

```

## PBPK-informed power versus animals per group

The two-sided test compares independent observed and PBPK-predicted log-AUC
means. Both groups are sampled from the **sparse-AUC empirical distribution**,
which includes drug/target IIV, FcRn-expression IIV, sparse sampling, and
observation error. The observed group is shifted by `log(2)`.

```{r}
#| label: power-functions

z_test_log_auc <- function(predicted_auc, observed_auc, alpha = ALPHA) {
  x <- log(predicted_auc)
  y <- log(observed_auc)
  se <- sqrt(var(x) / length(x) + var(y) / length(y))
  if (!is.finite(se) || se <= 0) return(FALSE)
  abs((mean(y) - mean(x)) / se) > qnorm(1 - alpha / 2)
}

simulate_single_dose_power <- function(predicted_reference,
                                       observed_reference,
                                       n,
                                       fold = EFFECT_FOLD,
                                       n_sim = N_POWER_SIM) {
  mean(replicate(n_sim, {
    predicted <- sample(predicted_reference, n, replace = TRUE)
    observed <- fold * sample(observed_reference, n, replace = TRUE)
    z_test_log_auc(predicted, observed)
  }))
}

power_curve <- crossing(
  Drug = MABS,
  Dose_mgkg = STUDY_DOSES,
  N_per_group = N_PER_GROUP
) %>%
  mutate(
    Power = pmap_dbl(list(Drug, Dose_mgkg, N_per_group), ~ {
      key <- reference_key(..1, ..2)
      predicted_reference <- rich_reference_list[[key]]
      observed_reference <- sparse_reference_list[[key]]
      simulate_single_dose_power(
        predicted_reference, observed_reference, ..3
      )
    }),
    # PBPK-predicted animals are virtual and therefore have zero animal cost.
    Total_animals = N_per_group,
    Power_per_animal = Power / Total_animals
  )

write.csv(
  power_curve,
  file.path(table_dir, "All_mAbs_power_curve.csv"),
  row.names = FALSE
)

power_curve %>%
  select(
    Drug, Dose_mgkg, N_per_group, Total_animals, Power, Power_per_animal
  ) %>%
  kable(
    digits = 3,
    caption = paste0(
      "Single-dose observed-versus-PBPK comparison. Total animals equals ",
      "n/group because the PBPK reference group is virtual. Multi-dose ",
      "animal accounting is presented in the next table."
    )
  ) %>%
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = FALSE)
```

## Two- versus three-dose-group designs

For each dose, an observed-versus-predicted log-AUC ratio and its variance are
calculated. Dose-specific ratios are combined by inverse-variance weighting.
The two-dose design uses 0.3 and 30 mg/kg; the three-dose design adds 3 mg/kg.

Scenario A fixes observed animals per dose group, so the three-dose design uses
more animals. Scenario B fixes the total number of observed animals and
allocates them as evenly as possible across dose groups. PBPK-predicted
reference samples are virtual and do not contribute to animal cost.

```{r}
#| label: dose-design-power

one_design_replicate <- function(drug, doses, n_by_dose,
                                 fold = EFFECT_FOLD) {
  estimates <- variances <- numeric(length(doses))

  for (i in seq_along(doses)) {
    key <- reference_key(drug, doses[i])
    predicted_reference <- rich_reference_list[[key]]
    observed_reference <- sparse_reference_list[[key]]
    n <- n_by_dose[i]
    predicted <- log(sample(predicted_reference, n, replace = TRUE))
    observed <- log(fold * sample(observed_reference, n, replace = TRUE))
    estimates[i] <- mean(observed) - mean(predicted)
    variances[i] <- var(observed) / n + var(predicted) / n
  }

  if (any(!is.finite(variances)) || any(variances <= 0)) return(FALSE)
  weights <- 1 / variances
  pooled_estimate <- weighted.mean(estimates, weights)
  pooled_se <- sqrt(1 / sum(weights))
  abs(pooled_estimate / pooled_se) > qnorm(1 - ALPHA / 2)
}

simulate_design_power <- function(drug, doses, n_by_dose,
                                  n_sim = N_POWER_SIM) {
  mean(replicate(
    n_sim,
    one_design_replicate(drug, doses, n_by_dose)
  ))
}

design_doses <- list(
  `2 dose groups` = c(0.3, 10),
  `3 dose groups` = c(0.3, 3, 10)
)

# A: fixed observed n in each dose group.
fixed_n_design <- crossing(
  Drug = MABS,
  Design = names(design_doses),
  N_per_group = N_PER_GROUP
) %>%
  mutate(
    Number_of_doses = lengths(design_doses[Design]),
    Total_animals = Number_of_doses * N_per_group,
    Power = pmap_dbl(list(Drug, Design, N_per_group), ~ {
      doses <- design_doses[[..2]]
      simulate_design_power(..1, doses, rep(..3, length(doses)))
    }),
    Comparison = "A: fixed n/group"
  )

# B: fixed total observed animals; require at least two animals per dose.
fixed_total_design <- crossing(
  Drug = MABS,
  Design = names(design_doses),
  Total_animals = FIXED_TOTAL_N
) %>%
  mutate(
    Number_of_doses = lengths(design_doses[Design]),
    Number_of_groups = Number_of_doses,
    N_per_group = floor(Total_animals / Number_of_groups),
    Animals_used = N_per_group * Number_of_groups
  ) %>%
  filter(
    N_per_group >= min(N_PER_GROUP),
    N_per_group <= max(N_PER_GROUP)
  ) %>%
  mutate(
    Power = pmap_dbl(list(Drug, Design, N_per_group), ~ {
      doses <- design_doses[[..2]]
      simulate_design_power(..1, doses, rep(..3, length(doses)))
    }),
    Comparison = "B: fixed total animals"
  )

design_power <- bind_rows(
  fixed_n_design %>%
    transmute(
      Drug, Design, Comparison, N_per_group, Total_animals,
      Animals_used = Total_animals, Power
    ),
  fixed_total_design %>%
    select(
      Drug, Design, Comparison, N_per_group, Total_animals,
      Animals_used, Power
    )
)

write.csv(
  design_power,
  file.path(table_dir, "All_mAbs_2_vs_3_dose_design_power.csv"),
  row.names = FALSE
)

fixed_n_design %>%
  transmute(
    Drug,
    Design,
    `Animals per dose group` = N_per_group,
    `Number of dose groups` = Number_of_doses,
    `Total animals` = Total_animals,
    Power,
    `Power per animal` = Power / Total_animals
  ) %>%
  arrange(Drug, `Animals per dose group`, `Number of dose groups`) %>%
  kable(
    digits = 3,
    caption = paste0(
      "Two- versus three-dose-group designs evaluated at n = 2-5 animals ",
      "per dose group. Total animals = n/group × number of dose groups."
    )
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  )

design_plot_a <- ggplot(
  fixed_n_design,
  aes(x = N_per_group, y = Power, color = Design)
) +
  geom_hline(yintercept = 0.8, linetype = 2, color = "grey35") +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_wrap(~ Drug, ncol = 4) +
  scale_color_viridis_d(
    option = "D", direction = -1, begin = 0.15, end = 0.80
  ) +
  scale_x_continuous(breaks = N_PER_GROUP) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  labs(
    x = "Observed animals per dose group",
    y = "Power",
    title = "A. Fixed animals per group",
    subtitle = "Total animal cost differs between designs",
    color = NULL
  ) +
  STUDY_THEME

design_plot_b <- ggplot(
  fixed_total_design,
  aes(x = Animals_used, y = Power, color = Design)
) +
  geom_hline(yintercept = 0.8, linetype = 2, color = "grey35") +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_wrap(~ Drug, ncol = 4) +
  scale_color_viridis_d(
    option = "D", direction = -1, begin = 0.15, end = 0.80
  ) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  labs(
    x = "Total animals used",
    y = "Power",
    title = "B. Fixed total animals",
    subtitle = "Observed animals allocated evenly across dose groups",
    color = NULL
  ) +
  STUDY_THEME

ggsave(
  file.path(figure_dir, "All_mAbs_dose_design_fixed_n.png"),
  design_plot_a, width = 6.5, height = 4.5, dpi = 300
)
ggsave(
  file.path(figure_dir, "All_mAbs_dose_design_fixed_total.png"),
  design_plot_b, width = 8, height = 5.5, dpi = 300
)

design_plot_a
design_plot_b
```

## Reproducibility record

```{r}
#| label: session-info

sessionInfo()
```
