---
title: "Process Simulation Outputs"
author: "NAMLink"
format:
  html:
    code-fold: true
    code-summary: "Show code"
    self-contained: true
---

# Set up

```{r}
#| label: setup
#| echo: true
#| warning: false
#| message: false

rm(list = ls())

library(tidyverse)
library(data.table)   # fast I/O (fread) and fast grouped operations
library(scales)
library(knitr)
library(kableExtra)
library(openxlsx)

source("00_functions.R")

datadir  <- "../Data/"
results_dir <- "../Results/SimulationResults"
figures_dir <- "../Results/Figures"

# AUC integration window: 28 days in hours
AUC_END_H <- 28 * 24   # 672 h
```

# Load data

## Observed data

```{r}
#| label: load-obs
#| message: false

obs_raw <- read.csv(
  file.path(datadir, "mAb_data_1.csv"),
  stringsAsFactors = FALSE,
  check.names      = FALSE
)

# Normalise species labels to match simulation scenario naming convention,
# convert time to hours, and keep only quantifiable rows
obs <- obs_raw %>%
  mutate(
    Species   = normalize_species(Species),
    Time_h    = case_when(
      tolower(trimws(`Time unit`)) %in% c("h", "hr", "hour", "hours") ~ as.numeric(Time),
      TRUE ~ as.numeric(Time) / 60
    ),
    Conc_ugmL = as.numeric(Measurement)   # already in µg/mL
  ) %>%
  filter(!is.na(Conc_ugmL), Conc_ugmL > 0)

# MW lookup: used for unit conversion of simulation output (µmol/L → µg/mL)
mw_lookup <- obs_raw %>%
  distinct(Molecule, MW = `Molecular Weight`) %>%
  mutate(MW = as.numeric(MW))

message("Observed data: ", nrow(obs), " rows | ",
        n_distinct(obs$Molecule), " molecules | ",
        n_distinct(obs$Species),  " species")
```

## Simulation results — latest run

```{r}
#| label: load-sims
#| message: false

# ── Auto-select the most recently modified results sub-folder ─────────────
sim_folders <- list.dirs(results_dir, recursive = FALSE, full.names = TRUE)
if (length(sim_folders) == 0) stop("No results folders found in: ", results_dir)

latest_dir <- sim_folders[which.max(file.mtime(sim_folders))]
message("Results folder: ", latest_dir)

# ── List scenario CSVs (exclude _population parameter files) ──────────────
csv_files <- list.files(latest_dir, pattern = "\\.csv$", full.names = TRUE)
csv_files <- csv_files[!grepl("_population\\.csv$", csv_files)]
message(length(csv_files), " scenario CSV files found")

# ── Helper: parse Molecule / Species / Dose from filename ─────────────────
# Convention: {Molecule}_{Species}_{Dose}_mpk.csv
parse_scenario <- function(path) {
  base  <- tools::file_path_sans_ext(basename(path))
  base  <- sub("_mpk$", "", base)
  parts <- strsplit(base, "_")[[1]]
  n     <- length(parts)
  list(
    Molecule = paste(parts[seq_len(n - 2)], collapse = "_"),
    Species  = parts[n - 1],
    Dose     = suppressWarnings(as.numeric(parts[n]))
  )
}

# ── Read CSVs in parallel (overlaps Google Drive I/O latency per file) ────
# Expected columns: IndividualId | Time [min] | {OutputPath} [µmol/l]
# Unit conversion: µmol/L × MW (g/mol) / 1000 = µg/mL
n_cores   <- max(1L, parallel::detectCores(logical = FALSE) - 1L)
n_workers <- min(n_cores, length(csv_files))
cl <- parallel::makeCluster(n_workers)
parallel::clusterEvalQ(cl, library(data.table))
parallel::clusterExport(cl, c("parse_scenario", "mw_lookup"), envir = environment())

sim_list <- parallel::parLapply(cl, csv_files, function(f) {
  meta <- parse_scenario(f)

  df <- tryCatch(
    data.table::fread(f, data.table = TRUE),
    error = function(e) NULL
  )
  if (is.null(df)) { message("Could not read: ", basename(f)); return(NULL) }

  # Find the plasma concentration column
  conc_col <- grep("Plasma \\(Peripheral Venous Blood\\)", names(df), value = TRUE)
  if (length(conc_col) == 0) return(NULL)

  mw_val <- mw_lookup$MW[mw_lookup$Molecule == meta$Molecule]
  if (length(mw_val) == 0 || is.na(mw_val[1])) {
    warning("MW not found for: ", meta$Molecule, " — skipping")
    return(NULL)
  }

  df[, .(
    Molecule     = meta$Molecule,
    Species      = meta$Species,
    Dose         = meta$Dose,
    IndividualId = as.integer(IndividualId),
    Time_h       = as.numeric(`Time [min]`) / 60,
    Conc_ugmL    = as.numeric(.SD[[conc_col[1]]]) * mw_val[1] / 1000
  ), .SDcols = conc_col[1]]
})

parallel::stopCluster(cl)
rm(cl)

# rbindlist is faster than dplyr::bind_rows for data.tables
sim <- data.table::rbindlist(Filter(Negate(is.null), sim_list))

# ── Clear diagnostic if output paths are not wired to the model ──────────
if (nrow(sim) == 0) {
  stop(
    "No concentration data found in any scenario CSV.\n\n",
    "Likely cause: the OutputPath in Scenarios.xlsx references the molecule\n",
    "name (e.g. 'Atezolizumab') but the .pkml model uses a different name internally.\n\n",
    "Fix: update the OutputPaths sheet to use the correct path and re-run 02_Run_Sims.qmd."
  )
}

n_scen <- sim[, .N, by = .(Molecule, Species, Dose)][, .N]
message("Loaded ", n_scen, " scenarios | ",
        uniqueN(sim$IndividualId), " individuals per scenario")
```

# Summarise for plotting

```{r}
#| label: summarise
#| message: false
#| warning: false

# ── Simulated: 5th / 50th / 95th percentile per time point ────────────────
# data.table grouped operation is 5–20× faster than dplyr on large tables
sim_summary <- sim[, .(
  Pred_p05 = quantile(Conc_ugmL, 0.05, na.rm = TRUE),
  Pred_med = quantile(Conc_ugmL, 0.50, na.rm = TRUE),
  Pred_p95 = quantile(Conc_ugmL, 0.95, na.rm = TRUE)
), by = .(Molecule, Species, Dose, Time_h)]

# ── Observed: mean ± SD per group / time point ────────────────────────────
obs_summary <- obs %>%
  group_by(Molecule, Species, Dose, Time_h) %>%
  summarise(
    Obs_mean = mean(Conc_ugmL, na.rm = TRUE),
    Obs_sd   = sd(Conc_ugmL,   na.rm = TRUE),
    n        = n(),
    .groups  = "drop"
  ) %>%
  mutate(
    Obs_lo = pmax(Obs_mean - Obs_sd, 1e-4),  # floor at small positive for log scale
    Obs_hi = Obs_mean + Obs_sd
  )
```

# PK profiles: observed vs simulated

One figure per molecule; panels = species; colour = dose.

```{r}
#| label: pk-plots
#| fig-width: 10
#| fig-height: 4.5
#| message: false

molecules <- sort(unique(c(sim_summary$Molecule, obs_summary$Molecule)))

pk_plots <- lapply(molecules, function(mol) {

  sim_mol <- sim_summary[Molecule == mol]   # data.table subset
  obs_mol <- filter(obs_summary, Molecule == mol)

  p <- ggplot() +

    # Prediction uncertainty ribbon (5th–95th percentile)
    geom_ribbon(
      data  = sim_mol,
      aes(x = Time_h, ymin = Pred_p05, ymax = Pred_p95, fill = factor(Dose)),
      alpha = 0.18
    ) +

    # Predicted median line
    geom_line(
      data      = sim_mol,
      aes(x = Time_h, y = Pred_med, colour = factor(Dose)),
      linewidth = 0.9
    ) +

    # Observed mean points
    geom_point(
      data  = obs_mol,
      aes(x = Time_h, y = Obs_mean, colour = factor(Dose)),
      shape = 16, size = 2.3
    ) +

    # Observed ± SD error bars
    geom_errorbar(
      data      = filter(obs_mol, !is.na(Obs_sd) & n > 1),
      aes(x = Time_h, ymin = Obs_lo, ymax = Obs_hi, colour = factor(Dose)),
      width     = 0.8, linewidth = 0.45
    ) +

    facet_wrap(~ Species, scales = "free_x") +
    scale_y_log10(limits = c(0.1, NA), 
      labels = label_number(drop0trailing = TRUE),
      breaks = 10^seq(-2, 4)
    ) +
    scale_colour_viridis_d(name = "Dose (mg/kg)", option = "D", end = 0.88) +
    scale_fill_viridis_d(name   = "Dose (mg/kg)", option = "D", end = 0.88) +
    labs(
      title    = mol,
      subtitle = "Lines = simulated median  |  Shading = 5th–95th percentile  |  Points = observed mean ± SD",
      x        = "Time (h)",
      y        = "Plasma Concentration (µg/mL)"
    ) +
    theme_bw(base_size = 12) +
    theme(
      strip.background = element_rect(fill = "grey93"),
      strip.text       = element_text(face = "bold"),
      legend.position  = "bottom",
      panel.grid.minor = element_blank()
    )

  ggsave(
    file.path(figures_dir, paste0("PK_ObsPred_", mol, ".png")),
    p, width = 10, height = 4.5, dpi = 150
  )
  message("Saved plot: ", mol)
  p
})

names(pk_plots) <- molecules
pk_plots
```

# PK metrics: AUC~0–28d~ and C~max~

AUC is integrated by the linear-log trapezoidal rule over 0 –
`r AUC_END_H` h (28 days). For observed subjects the integral is
truncated at their last measured time point when that falls before 672 h.
Fold error = Predicted / Observed; values within 0.5–2.0 (2-fold) are
highlighted green.

```{r}
#| label: pk-metrics
#| message: false
#| warning: false

# ── Vectorised linear-log trapezoidal AUC ─────────────────────────────────
# Replaces the previous for-loop with vectorised diff/ifelse — ~100× faster
# when called across thousands of individual-level groups.
trapz_auc <- function(time, conc, t_end = Inf) {
  keep <- !is.na(time) & !is.na(conc) & conc > 0 & time <= t_end
  t <- time[keep]
  c <- conc[keep]
  if (length(t) < 2) return(NA_real_)

  dt  <- diff(t)
  c1  <- c[-length(c)]
  c2  <- c[-1]

  # Use log trapezoid on strictly declining segments; linear elsewhere
  use_log <- c1 != c2 & c1 > 0 & c2 > 0
  trap <- ifelse(
    use_log,
    dt * (c1 - c2) / log(c1 / c2),   # log trapezoid
    dt * (c1 + c2) / 2                # linear trapezoid
  )
  sum(trap)
}

# ── Simulated metrics: fully vectorised — no per-group R function calls ────
# Sort once, then compute lead/lag differences across all groups simultaneously.
# This replaces thousands of trapz_auc() dispatches with three data.table ops.
setorder(sim, Molecule, Species, Dose, IndividualId, Time_h)

sim[, c("dt_h", "c_lag") := .(
  Time_h    - shift(Time_h),
  shift(Conc_ugmL)
), by = .(Molecule, Species, Dose, IndividualId)]

# Linear-log trapezoid rule: log branch where both ends > 0 and differ
sim[, trap := fifelse(
  is.na(dt_h) | dt_h <= 0 | is.na(c_lag) | c_lag <= 0 | Conc_ugmL <= 0, 0,
  fifelse(
    c_lag != Conc_ugmL,
    dt_h * (c_lag - Conc_ugmL) / log(c_lag / Conc_ugmL),
    dt_h * (c_lag + Conc_ugmL) / 2
  )
)]

sim_metrics <- sim[Time_h <= AUC_END_H, .(
  Pred_AUC  = sum(trap,         na.rm = TRUE),
  Pred_Cmax = max(Conc_ugmL,   na.rm = TRUE)
), by = .(Molecule, Species, Dose, IndividualId)][, .(
  Pred_AUC  = mean(Pred_AUC,  na.rm = TRUE),
  Pred_Cmax = mean(Pred_Cmax, na.rm = TRUE)
), by = .(Molecule, Species, Dose)]

sim[, c("dt_h", "c_lag", "trap") := NULL]

# ── Observed metrics: per subject → group mean ────────────────────────────
obs_metrics <- obs %>%
  group_by(Molecule, Species, Dose, `Subject Id`) %>%
  summarise(
    Obs_AUC  = trapz_auc(Time_h, Conc_ugmL, t_end = AUC_END_H),
    Obs_Cmax = max(Conc_ugmL, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  group_by(Molecule, Species, Dose) %>%
  summarise(
    Obs_Cmax = mean(Obs_Cmax, na.rm = TRUE),
    Obs_AUC  = mean(Obs_AUC,  na.rm = TRUE),
    .groups  = "drop"
  )

# ── Join and compute fold errors ──────────────────────────────────────────
metrics_tbl <- full_join(
    as.data.frame(obs_metrics),
    as.data.frame(sim_metrics),
    by = c("Molecule", "Species", "Dose")
  ) %>%
  mutate(
    FE_Cmax = round(Pred_Cmax / Obs_Cmax, 2),
    FE_AUC  = round(Pred_AUC  / Obs_AUC,  2)
  ) %>%
  arrange(Molecule, Species, Dose)
```

## Human

```{r}
#| label: metrics-table-human

# Colour-code fold error cells: green = within 2-fold, red = outside
fmt_fe <- function(x) {
  cell_spec(
    formatC(x, format = "f", digits = 2),
    background = ifelse(!is.na(x) & x >= 0.5 & x <= 2.0, "#d4edda", "#f8d7da"),
    color      = ifelse(!is.na(x) & x >= 0.5 & x <= 2.0, "#155724", "#721c24"),
    bold       = TRUE
  )
}

# Shared helper: format and render one species-level table
render_metrics_kbl <- function(df, caption) {
  df_fmt <- df %>%
    mutate(
      `Obs Cmax (µg/mL)`   = round(Obs_Cmax, 1),
      `Pred Cmax (µg/mL)`  = round(Pred_Cmax, 1),
      `FE Cmax`            = map_chr(FE_Cmax, fmt_fe),
      `Obs AUC (µg·h/mL)`  = round(Obs_AUC,  0),
      `Pred AUC (µg·h/mL)` = round(Pred_AUC, 0),
      `FE AUC`             = map_chr(FE_AUC,  fmt_fe)
    ) %>%
    select(
      Molecule, `Dose (mg/kg)` = Dose,
      `Obs Cmax (µg/mL)`, `Pred Cmax (µg/mL)`, `FE Cmax`,
      `Obs AUC (µg·h/mL)`, `Pred AUC (µg·h/mL)`, `FE AUC`
    )

  df_fmt %>%
    kbl(
      escape  = FALSE,
      align   = c("l", "r", "r", "r", "c", "r", "r", "c"),
      caption = caption
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE, font_size = 13
    ) %>%
    column_spec(5, width = "5em") %>%
    column_spec(8, width = "5em") %>%
    pack_rows(index = setNames(
      rle(df$Molecule)$lengths,
      rle(df$Molecule)$values
    ))
}

caption_suffix <- paste0("AUC integrated over 0–", AUC_END_H, " h (28 days). FE = Predicted / Observed.")

metrics_tbl %>%
  filter(Species == "Human") %>%
  render_metrics_kbl(paste("Human PK Metrics —", caption_suffix))
```

## NHP

```{r}
#| label: metrics-table-nhp

metrics_tbl %>%
  filter(Species == "Monkey") %>%
  render_metrics_kbl(paste("NHP PK Metrics —", caption_suffix))
```

## Export to Excel

```{r}
#| label: save-metrics
#| message: false
#| output: false

metrics_out <- metrics_tbl %>%
  rename(
    `Dose (mg/kg)`       = Dose,
    `Obs Cmax (µg/mL)`   = Obs_Cmax,
    `Pred Cmax (µg/mL)`  = Pred_Cmax,
    `Fold Error Cmax`    = FE_Cmax,
    `Obs AUC (µg·h/mL)`  = Obs_AUC,
    `Pred AUC (µg·h/mL)` = Pred_AUC,
    `Fold Error AUC`     = FE_AUC
  ) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

wb <- createWorkbook()
addWorksheet(wb, "PK_Metrics")
writeData(wb, "PK_Metrics", metrics_out)
setColWidths(wb, "PK_Metrics",
             cols   = seq_len(ncol(metrics_out)),
             widths = c(16, 10, 13, 18, 18, 15, 18, 18, 15))

out_xlsx <- file.path(figures_dir, "PK_Metrics_ObsPred.xlsx")
saveWorkbook(wb, out_xlsx, overwrite = TRUE)
message("Saved: ", out_xlsx)
```
